---
title: "Practice exam - Apartment applicants"
output:
  html_document:
    code_folding: show
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
date: "`r format(Sys.time(), '%d %B, %Y')`"
subtitle: ExamPA.net
---

```{r warning = F, message=F}
#NO CHANGES NEEDED
set.seed(1)
library(knitr)
library(broom)
library(ggplot2)
library(plyr)
library(corrplot)
library(caret)
library(gridExtra)
library(scales)
library(Rmisc)
library(randomForest)
library(psych)
library(xgboost)
library(rpart)
library(rpart.plot)
library(randomForest)
library(tidyverse)
library(ExamPAData)
theme_set(theme_minimal())

df <- apartment_apps


```

## Set up

Your assistant has provided you with these code templates.

```{r eval = F}
#This function will calculate the log-liklihood based on a Poisson model for the number of applicants
LLfunction <- function(targets, predicted_values){
  p_v_zero <- ifelse(predicted_values <= 0, 0, predicted_values)
  p_v_pos <- ifelse(predicted_values <= 0, 0.000001 ,predicted_values)
  return(sum(targets*log(p_v_pos)) - sum(p_v_zero))
}
print("loglikelihood")
LLfunction(test$applicants,predictions)
```


This code creates a scatter plot, a box plot, and a histogram.


```{r eval = F}
#create a boxplot
ggplot(df, aes(as.factor(applicants),VARIABLE)) + 
  geom_boxplot()

#create a scatterplot
ggplot(df,aes(applicants,VARIABLE)) + 
  geom_point()

#create a histogram
ggplot(df,aes(VARIABLE)) + 
  geom_histogram()

#create a bar plot
ggplot(df, aes(VARIABLE)) + 
  geom_bar(stat = "count")
```

Shows the average number of applicants across factor levels.

```{r eval = F}
#Average number of health applicants per unit
df %>% 
  group_by(VARIABLE) %>% 
  summarise(
    average_num_applicants = sum(applicants*num_units)/sum(num_units)
  )
```

Converts variables to numeric or factor while also setting the base reference level to the value which has the most observations.

```{r eval = F}
##Convert to factor and set factor levels
df <- df %>% mutate(VARIABLE = fct_infreq(as.character(VARIABLE))

#Using Base R
df$VARIABLE = fct_infreq(as.character(df$VARIABLE))

#Convert to numeric
df <- df %>% mutate(VARIABLE = fct_infreq(as.characterVARIABLE))

#Using Base R
df$VARIABLE = as.numeric(df$VARIABLE)
```


## Task 1 - Examine the target variable and number of units

```{r}
glimpse(df)
summary(df)
sum(df$num_units)
sum(df$num_units*df$applicants)/sum(df$num_units)
df %>% summarise(avg_applicants = sum(applicants*num_units)/sum(num_units))


#NO CODE CHANGES NEEDED
df %>% 
  group_by(applicants) %>% 
  summarise(total_units = sum(num_units))

#Graph A: Histogram
ggplot(data=df, aes(applicants)) +
        geom_histogram()

#Graph B: Bar plot
df %>% 
  mutate(applicants = as.factor(applicants)) %>% 
  group_by(applicants) %>% 
  summarise(
    total_units = sum(num_units)
  ) %>% 
  ungroup() %>% 
  ggplot(aes(x=applicants, y=total_units)) + 
  geom_bar(stat="identity") 
  
```

## Task 2 - Explore the predictor variables

```{r}
summary(df)
#NO CHANGES NEEDED
#Calculate weighted averages with num_units as the weights
df %>% 
  summarise(
    mean_number_of_units = sum(num_units),
    mean_sale_price = dollar(sum(num_units*sale_price)/sum(num_units)),
    mean_overall_qual = sum(num_units*overall_qual)/sum(num_units)
  )

#neighborhood sale price
df %>% arrange(desc(neighborhood_saleprice)) %>% head()


#Sale Price
df %>% 
  ggplot(aes(sale_price,y=..density.., weight = num_units)) + 
  geom_histogram()


df %>% 
  ggplot(aes(sale_price/neighborhood_saleprice,y=..density.., weight = num_units)) + 
  geom_histogram()

df %>% 
  ggplot(aes(neighborhood_saleprice)) + 
  geom_histogram()

df <- df %>% mutate(log_rel_price = log(sale_price/neighborhood_saleprice),
                    log_sale_price = log(sale_price)) %>% 
  dplyr::select(-sale_price, -neighborhood_saleprice)
```


```{r}
#overall qual
ggplot(df,aes(overall_qual)) + 
  geom_bar(stat = "count")

df %>% 
  summarise(
    avg = sum(num_units*overall_qual)/sum(num_units)
  ) 

df %>% 
  group_by(overall_qual) %>% 
  summarise(
    total_units = sum(num_units)
  ) %>% 
  ggplot(aes(x=overall_qual, y=total_units)) + 
  geom_bar(stat="identity") 

#central air

df %>% 
  group_by(central_air) %>% 
  summarise(
    total_units = sum(num_units)
  ) %>% 
  ggplot(aes(x=central_air, y=total_units)) + 
  geom_bar(stat="identity") 

df %>% 
  group_by(central_air) %>% 
  summarise(
    total_units = sum(num_units)
  ) 

df %>% count(overall_qual)

#check that bae level (the one that appears first) is the one with the most observations
summary(df$central_air)
```

```{r}
df %>% group_by(central_air) %>% 
  summarise(applicants=sum(applicants),
            total_units = sum(num_units))

ggplot(df,aes(as.factor(central_air),applicants)) + 
  geom_boxplot()

df$central_air <- fct_infreq(df$central_air)
summary(df$central_air)
```

## Task 3 - Engineer three additional features

Code is provided to calculate the average and create box plots of sale_price by year and month.  

```{r}
df <- df %>% 
  mutate(
    sale_price_per_sqft = exp(log_rel_price) / total_sq_feet,
    bath_pr_sqft = tot_bathrooms/total_sq_feet,
    student_apt = ifelse(month_sold %in% c(7,8),1,0)
  )
df %>% count(student_apt)

df %>% summary()
```


## Task 4 - Inspect the garage_type variables

No code is provided.

```{r}
df %>% filter(
  (garage_type_attchd + garage_type_basment + garage_type_detchd + garage_type_builtIn + garage_type_no_garage) > 0)

#rows_with_errors <- (df$garage_type_attchd==0) + (df$garage_type_detchd==0) + (df$garage_type_builtIn==0) + (df$garage_type_no_garage==0)

#df <- df[-rows_with_errors]
```


No changes are needed to the below code.  It is included because the results may be helpful for subsequent tasks.

```{r}
#NO CHANGES NEEDED
#Example: the 18th property has 0's for all garage type variables
df %>% dplyr::select(contains("garage")) %>%  dplyr::slice(18)

#NO CHANGES NEEDED
#Calculate the Number of Units for each garage type
df %>% dplyr::select(num_units,contains("garage")) %>% gather(feature,value,-num_units) %>% group_by(feature) %>%  summarise(total_units=sum(num_units*value))

#NO CHANGES NEEDED
#Calculate the Number of Units for each neighborhood
df %>% dplyr::select(num_units,contains("neighborhood")) %>% gather(feature,value,-num_units) %>% group_by(feature) %>%  summarise(total_units=sum(num_units*value))
```

You do not need to check the neighborhood columns as your assistent has already verified that these are correct for all properties.

## Task 5 - Select GLM parameters

No code is provided

## Task 6 - Fit a GLM

```{r}
#NO CHANGES NEEDED
#Create training and test sets
index <- createDataPartition(y = df$applicants, p = 0.8, list = F)

train <- df %>% dplyr::slice(index)
test <- df %>% dplyr::slice(-index)

train_x <- train[,-1]
train_y <- train$applicants

#standardize to be between 0 and 1
train_number_of_units <- train$num_units/sum(train$num_units)

test_x <- test[,-1]
test_y <- test$applicants

#standardize to be between 0 and 1
test_number_of_units <- test$num_units/sum(test$num_units)
```



```{r}
#the base (reference) level for neighborhood is the one with the most observations
#the base (reference) level for garage is the one with the most observations

#Fit a GLM
#Do not use the offset or weight variables as predictors.  Remove them from the formula (as the below is doing)
#Type ?family into the R console to see options for FAMILY
glm <- glm(
  applicants ~ . + offset(log(num_units)) - neighborhood_n_ames - garage_type_attchd - num_units - log_sale_price, 
  data = train,
  family = poisson(link = "log")
)
AIC(glm)
summary(glm)
```


```{r}
#Alternative answer
#Models claim frequency (average applicants per unit)
glm2 <- glm(
  applicants/num_units ~ . + - neighborhood_n_ames - garage_type_attchd  - log_sale_price - num_units, 
  data = train,
  weights = num_units,
  family = poisson(link = "log")
)
AIC(glm2)
summary(glm2)

#You can verify that both models return the same predictions

app_count <- predict(glm, type="response", newdata = test)

app_frequency <- predict(glm2, type="response", newdata = test)*test$num_units

head(app_count)
head(app_frequency)
```


## Task 7 - Use AIC to select features

```{r}
#NO CODE CHANGES NEEDED
#The following code will use the same formula and family from the `glm` object to perform stepwise selection

library(MASS)
stepwise_result <- stepAIC(glm)
library(dplyr)

final_glm <- glm(
  stepwise_result$formula, 
  data = train,
  family = glm$family
)

AIC(final_glm)
summary(final_glm)
plot(final_glm)
```


## Task 8 - Fit a LASSO


```{r}
#THE ONLY CHANGES NEEDED
#1. Replace "WEIGHT_VARIABLE" with the weight variable that you used for the GLM
#2. In the  "lasso_formula" below, use your formula from task 6
#3. Use the same FAMILY from task 6.  Type ?glmnet into the console to read the possible options

lasso_formula <- applicants ~ . + offset(log(num_units)) - neighborhood_n_ames - garage_type_attchd - num_units - log_sale_price 

lasso_x_train <- model.matrix(lasso_formula, data = train)
lasso_x_test <- model.matrix(lasso_formula, data = test)

control <-trainControl(method="cv", number=15)
lassoGrid <- expand.grid(alpha = 1, lambda = seq(0.0001,0.01,by = 0.0005))

#Fit a lasso and inspect the variables which have zero coefficients.
#Note that the glmnet library only allows the identity link function
lasso <- train(x = lasso_x_train, 
               y = train_y, 
               method = 'glmnet', 
               family = "poisson",

               trControl = control, 
               tuneGrid = lassoGrid) 

#This code gets the coefficients from the LASSO which are zero
lasso_results <- varImp(lasso,scale=F)
lasso_coefficients <- lasso_results$importance

#Show the coefficients which are zero
variables_with_zeros <- colnames(lasso_x_train)[which(lasso_coefficients$Overall==0)]
print("Variables with Coefficients of Zero:")
cat( paste( variables_with_zeros, collapse='\n' ) )
```

## Task 9 - Create a bagged tree model

The following code creates eight samples of the data.  20% of records are taken, with replacement, from the data randomly.  

```{r fig.width=10, fig.height=10}
sample1 <- train %>% sample_frac(0.2,replace=T)
sample2 <- train %>% sample_frac(0.2,replace=T)
sample3 <- train %>% sample_frac(0.2,replace=T)
sample4 <- train %>% sample_frac(0.2,replace=T)
sample5 <- train %>% sample_frac(0.2,replace=T)
sample6 <- train %>% sample_frac(0.2,replace=T)
sample7 <- train %>% sample_frac(0.2,replace=T)
sample8 <- train %>% sample_frac(0.2,replace=T)
```


This is the setting for the decision tree paramters.  Make your adjustments here to test out different paramters.

```{r}
#ctrl <- rpart.control(minbucket = 5, cp = 0.01, maxdepth = 5)
ctrl <- rpart.control(minbucket = 10, cp = 0.001, maxdepth = 5)
#ctrl <- rpart.control(minbucket = 10, cp = 0.01, maxdepth = 7)
```


The following code sets up a decision tree using all the variables in the dataframe.  Each tree uses only 20% of the records.  The number of units are used as the weights.

```{r warning = F, fig.height=8, fig.width=8}
tree1 <- rpart(data = sample1, lasso_formula, weights = sample1$num_units, control = ctrl)
tree2 <- rpart(data = sample2, lasso_formula, weights = sample2$num_units, control = ctrl)
tree3 <- rpart(data = sample3, lasso_formula, weights = sample3$num_units, control = ctrl)
tree4 <- rpart(data = sample4, lasso_formula, weights = sample4$num_units, control = ctrl)
tree5 <- rpart(data = sample5, lasso_formula, weights = sample5$num_units, control = ctrl)
tree6 <- rpart(data = sample6, lasso_formula, weights = sample6$num_units, control = ctrl)
tree7 <- rpart(data = sample7, lasso_formula, weights = sample7$num_units, control = ctrl)
tree8 <- rpart(data = sample8, lasso_formula, weights = sample8$num_units, control = ctrl)

rpart.plot(tree1, cex = .7)
rpart.plot(tree2, cex = .7)
rpart.plot(tree3, cex = .7)
rpart.plot(tree4, cex = .7)
rpart.plot(tree5, cex = .7)
rpart.plot(tree6, cex = .7)
rpart.plot(tree7, cex = .7)
rpart.plot(tree8, cex = .7)
```

```{r}
#make predictions
tree1_pred <- predict(tree1, newdata = test, type = "vector")
tree2_pred <- predict(tree2, newdata = test, type = "vector")
tree3_pred <- predict(tree3, newdata = test, type = "vector")
tree4_pred <- predict(tree4, newdata = test, type = "vector")
tree5_pred <- predict(tree5, newdata = test, type = "vector")
tree6_pred <- predict(tree6, newdata = test, type = "vector")
tree7_pred <- predict(tree7, newdata = test, type = "vector")
tree8_pred <- predict(tree8, newdata = test, type = "vector")

#UPDATE THIS FORMULA TO PERFORM BAGGING
trees_pred <- (tree1_pred + tree2_pred + tree3_pred+ tree4_pred+ tree5_pred+ tree6_pred+ tree7_pred+ tree8_pred)/8

LLfunction <- function(targets, predicted_values){
  p_v_zero <- ifelse(predicted_values <= 0, 0, predicted_values)
  p_v_pos <- ifelse(predicted_values <= 0, 0.000001 ,predicted_values)
  return(sum(targets*log(p_v_pos)) - sum(p_v_zero))
}
# "targets" is a vector containing the actual values for the target variable
# "predicted_values" is a vector containing the predicted values for the target variable

LLfunction(test$applicants, trees_pred)

LLfunction(test$applicants, tree1_pred)
```


## Task 10 - Measure the variable importance with a Random Forest

```{r}
#NO CHANGES NEEDED
RF <- randomForest( applicants ~ . + neighborhood_n_ames - garage_type_attchd - num_units - log_sale_price,
                   data = train,
                   weights = train_number_of_units, 
                   ntree=400,
                   importance=TRUE)

imp_RF <- importance(RF)
imp_DF <- data.frame(Variables = row.names(imp_RF), MSE = imp_RF[,1])
imp_DF <- imp_DF[order(imp_DF$MSE, decreasing = TRUE),]

ggplot(imp_DF[1:30,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none")
```

## Task 11 - Compare model performance

```{r}
#This function will calculate the log-liklihood based on a Poisson model for the number of applicants
LLfunction <- function(targets, predicted_values){
  p_v_zero <- ifelse(predicted_values <= 0, 0, predicted_values)
  p_v_pos <- ifelse(predicted_values <= 0, 0.000001, predicted_values)
  return(sum(targets*log(p_v_pos)) - sum(p_v_zero))
}
print("loglikelihood")

glm_pred <- predict(final_glm, type="response", newdata = test)
lasso_pred <- exp(predict(lasso,newdata = lasso_x_test, weights = test_number_of_units))
RF_pred <- predict(RF, type="response", newdata = test)

tibble(Model = c("GLM", "LASSO", "Bagged Trees", "Random Forest"),
       LogLikelihood = c(LLfunction(test$applicants,glm_pred),
                         LLfunction(test$applicants,lasso_pred),
                         LLfunction(test$applicants, trees_pred),
                         LLfunction(test$applicants,RF_pred)))
```

## Task 12 - Executive Summary

