---
title: "Data Loans - Solution"
output:
  html_document:
    code_folding: show
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
subtitle: ExamPA.net
---

```{r message = F, warning = F}
library(plyr)
library(tidyverse)
library(scales)
library(ExamPAData)
theme_set(theme_bw())
df <- bank_loans
```

## Task 1 - Examine the Target Varible

The following can be used 

```{r eval = F}
# Remove column
df$varname <- NULL

# I convert character VARIABLE to a factor.
df$admit_type_id <- as.factor(df$admit_type_id)

# Exclude certain values
df <- subset(df, VARIABLE != "VALUE") # Replace VARIABLE and VALUE
# Value is in quotation marks for factor variables. No quotation marks for numeric variables.
```

Combine variable levels

```{r eval = F}
#Using the tidyverse (fastest method)
df <- df %>% mutate(VARIABLE = ifelse(VARIABLE %in% c("VALUE1", "VALUE2", "VALUE3"), "other", VARIABLE))

# Replace VARIABLE with the variable name to have reduced number of levels.
# Replace LEVELs with new level names.

print("Data Before Combine Levels")
table(df$VARIABLE) # Replace VARIABLE

# Combine levels of VARIABLE by mapping one level to another level

var.levels <- levels(as.factor(df$VARIABLE)) # Replace VARIABLE
df$VARIABLE <- mapvalues(df$VARIABLE, var.levels,
		c("basic.4y", "basic.6y", "basic.9y", "high.school", "other", "professional.course", "university.degree", "other")) # Replace VARIABLE twice and replace LEVELs with the new names.

print("Data After Combine Levels")
table(df$VARIABLE) # Replace VARIABLE

# rm(var.levels)
```

Remove records

```{r eval = F}
#Remove outliers
df <- df %>% filter(VARIABLE >= VALUE)
#in Base R
#df <- subset(df, VARIABLE >= 40)
#Can also use != or <=
```


This code will show the percentage of `y` and number of records across different levels of a categorical VARIABLE.

```{r eval = F}
df %>% group_by(VARIABLE) %>% 
  summarise(percent_subscribed = percent(mean(y=="yes"), accuracy = 0.01),
            number_of_policies = n()) %>% 
  arrange(number_of_policies)
```



```{r}
glimpse(df)
```

```{r}
df %>% count(y)
```


```{r}
summary(df)

df %>% group_by(job) %>% 
  summarise(percent_subscribed = percent(mean(y=="yes"), accuracy = 0.01),
            number_of_policies = n()) %>% 
  arrange(number_of_policies)

df %>% group_by(marital) %>% 
  summarise(percent_subscribed = percent(mean(y=="yes"), accuracy = 0.01),
            number_of_policies = n()) %>% 
  arrange(number_of_policies)

df %>% group_by(education) %>% 
  summarise(percent_subscribed = percent(mean(y=="yes"), accuracy = 0.01),
            number_of_policies = n()) %>% 
  arrange(number_of_policies)

df %>% group_by(housing) %>% 
  summarise(percent_subscribed = percent(mean(y=="yes"), accuracy = 0.01),
            number_of_policies = n()) %>% 
  arrange(number_of_policies)

df %>% group_by(loan) %>% 
  summarise(percent_subscribed = percent(mean(y=="yes"), accuracy = 0.01),
            number_of_policies = n()) %>% 
  arrange(number_of_policies)

df %>% group_by(contact) %>% 
  summarise(percent_subscribed = percent(mean(y=="yes"), accuracy = 0.01),
            number_of_policies = n()) %>% 
  arrange(number_of_policies)

df %>% group_by(day_of_week) %>% 
  summarise(percent_subscribed = percent(mean(y=="yes"), accuracy = 0.01),
            number_of_policies = n()) %>% 
  arrange(number_of_policies)

df %>% group_by(poutcome) %>% 
  summarise(percent_subscribed = percent(mean(y=="yes"), accuracy = 0.01),
            number_of_policies = n()) %>% 
  arrange(number_of_policies)
```

Combine levels for education

```{r}


df <- df %>% mutate(education = ifelse(education %in% c("illiterate", "unknown"), "other", education))
df %>% count(education)
#using base R

# Replace VARIABLE with the variable name to have reduced number of levels.
# Replace LEVELs with new level names.
# 
# print("Data Before Combine Levels")
#table(df$education) # Replace VARIABLE
# 
# # Combine levels of VARIABLE by mapping one level to another level
# 
#var.levels <- levels(as.factor(df$education)) # Replace VARIABLE
#df$education <- mapvalues(df$education, var.levels,
# 		c("basic.4y", "basic.6y", "basic.9y", "high.school", "other", "professional.course", "university.degree", "other")) # Replace VARIABLE twice and replace LEVELs with the new names.
# 
# print("Data After Combine Levels")
# table(df$education) # Replace VARIABLE

# rm(var.levels)
```

## Task 2 - Decide on which variables to discard

```{r}
#the data dictionary says to not use duration in predictive modeling
df <- df %>% select(-duration)

#base R 
#df$duration <- NULL
```

## Task 3 - Examine the numeric variables

```{r}
df %>% select_if(is.numeric) %>% summary()
```


The following creates histograms of the numeric variables

```{r}
df %>% 
  select_if(is.numeric) %>% 
  gather(variable, value) %>% 
  ggplot(aes(value)) + 
  geom_histogram() + 
  facet_wrap(vars(variable), scales = "free")
```


```{r}
df %>% select_if(is.numeric) %>% summary()

#there are several records above the max of 40 in the data dictionary.  These were removed.
df %>% count(campaign)

#Remove outliers
df <- df %>% filter(campaign <= 40)
#in Base R
#df <- subset(df, campaign <= 40)

#example creating a histogram
df %>% ggplot(aes(cons.price.idx)) + geom_histogram()
df %>% count(emp.var.rate)

df %>% count(pdays)

#example of how to create bins
df <- df %>% 
  mutate(pdays_bin = case_when(pdays <= 5 ~ "0-5",
                               pdays <= 10 ~ "0-10",
                               pdays <= 30 ~ "10-30",
                               pdays == 999 ~ "None"))

#remove the original pdays variable
df <- df %>% select(-pdays)

df %>% count(pdays_bin)


# Releveling factor variables

df <- mutate_if(df, is.character, fct_infreq)

# Note: the SOA template code has a section for releveling factors, however, this frequently has errors if you have used any tidyverse functions on factor variable previously.  We recommend that you just remember the above lines of code so that you will not encounter these technical difficulties during the exam
glimpse(df)
```

## Task 4 - Examine the factor variables

The following will create bar charts for you. 

```{r}
#specify the categorical variables (factors) that you want to graph
#If you created any new variables, add them to this list
categorical_vars <- c("job", "marital", "education", "housing", "loan", "contact", "day_of_week", "poutcome","pdays_bin")
```

```{r}
#No changes are needed to this code chunk
library(ggpubr)

make_graphs <- function(i = "job"){
  temp_df <- df %>% select(i, "y")
  names(temp_df) <- c("x","y")
  
  fct_order <- temp_df %>% 
  count(x) %>% 
  arrange(n) %>% 
  select(x) %>% 
  unlist() %>% 
  as.character()

  p1 <- temp_df %>% 
    mutate(x = fct_relevel(x, fct_order)) %>% 
    ggplot(aes(x, fill = y)) + 
    geom_bar(stat = "count") + 
    coord_flip() + 
    theme(legend.position = "top") + 
    xlab(i) + 
    ylab("Total Number of People")
  
  p2 <- temp_df %>% 
    mutate(x = fct_relevel(x, fct_order)) %>% 
    ggplot(aes(x, fill = y)) + 
    geom_bar(stat = "count", position = "fill") + 
    coord_flip() + 
    theme(legend.position = "top") + 
    xlab(i) + 
    ylab("Percent of People")
  
  plot(ggarrange(p2, p1))
}

for(var in categorical_vars){
  make_graphs(var)
}
```
