---
title: "Customer Phone Calls -  Solution"
subtitle: "ExamPA.net Solution to SOA PA 6/18/20"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    theme: cerulean
    toc: true
    code_folding: show
    toc_float: true
---

For your convenience, this solution document has been formatted to be easy to read.  When taking this exam, candidates are not expected to make their `.Rmd` documents look this way.  The `make_table` function is included only to improve your reading experience.  The SOA's instructions from the Project Statement say only that

>At a minimum you must submit your completed report template and an Rmd file that supports your work.  Graders expect that your Rmd code can be run from beginning to end.  The code snippets provided should either be commented out or adapted for execution.  Ensure that it is clear where in the code each of the tasks is addressed.  Documentation of your thought process and conclusion for each task should be completely contained within your Word report.  The Rmd code should be clear, may contain commentary, and support your work.


# Task 1: Explore the data

```{r setup, message=F, warning=F}
#NO CHANGES NEEDED
# Load needed libraries.
library(ggplot2)
library(dplyr)
library(tidyr)
library(caret)
library(kableExtra) #If needed, install this package.  It's just used for making tables.
library(ExamPAData)

theme_set(theme_minimal())

#function to create html tables
make_table <- function(df, caption = ""){
  df %>% 
    kable(caption = caption) %>% 
    kable_styling(bootstrap_options = "striped", full_width = F)}

#global chunk options
knitr::opts_chunk$set(warning = F, message = F, echo = T, fig.width = 14, fig.height = 8)
```

```{r}
# Read the data.
df <- customer_phone_calls
  
summary(df)

# Create a histogram for irate.
ggplot(df, aes(x = irate)) +
  geom_histogram(binwidth = 5, fill = "royalblue", col = "royalblue")

# Create bar chart for month.
ggplot(df, aes(x = month)) +
  geom_bar(stat = "count", fill = "royalblue", col = "royalblue") +
  theme(axis.text = element_text(size = 6))

# Create a bar chart for edu_years.
ggplot(df, aes(x = edu_years)) +
  geom_bar(stat = "count", fill = "royalblue", col = "royalblue") +
  theme(axis.text = element_text(size = 6))

# Create a boxplot of the age distribution for different jobs.
boxplot(age ~ job, data = df, ylab = "Age Distribution", cex.axis = 0.5)

# 	Create a graph showing the proportion purchasing by age.
ggplot(df) +
  aes(x = age, fill = factor(purchase)) +
  labs(y = "Proportion of Purchases") +
  ggtitle("Proportion of Purchases by age") +
  geom_bar(position = "fill")

# Create a graph showing the proportion purchasing by month.
ggplot(df) +
  aes(x = month, fill = factor(purchase)) +
  labs(y = "Proportion of Purchases") +
  ggtitle("Proportion of Purchases by month") +
  geom_bar(position = "fill")
```

# Task 2: Consider the education variable.

No code provided.

```{r}
summary(df)
df %>% count(edu_years)
df %>% count(loan)

```


# Task 3: Handle missing values.


```{r}

# Check missing values. Display missing proportions for each variable that has them.
missing_proportion <- colMeans(is.na(df))
missing_data <- data.frame(colnames = colnames(df), missing_proportion = missing_proportion)
missing_data %>%
  filter(missing_proportion > 0) %>%
  ggplot(aes(x = colnames, y = missing_proportion, label = missing_proportion)) +
  geom_bar(stat = "identity", fill = "royalblue", col = "royalblue")

summary(df)

# The code below calculates the proportion of purchases for NAs and for nonNAs for each variable that has NAs.
#
print("Purchase Proportions by variable, for missing and non missing values")
print(sprintf("%10s %15s %15s", "Variable", "PP_for_NAs", "PP_for_non_NAs"))
varnames <- c("housing", "job", "loan", "marital", "edu_years")
for (t in varnames)
{
  ind <- is.na(df[t])
  print(sprintf("%10s %15.2f %15.2f", t, mean(df["purchase"][ind]), mean(df["purchase"][!ind])))
}
```

```{r}
#edu_years

# Impute using the mean (works only for numeric variables)
df$edu_years[is.na(df$edu_years)] <- mean(df$edu_years, na.rm = TRUE)
summary(df)
df %>% count(job)

df %>% count(marital)
summary(df)

#housing
# Remove rows
df <- df[!is.na(df$housing), ]

#job
# Convert missing values to "unknown" (works only for factor variables)

# First create a new level.
levels(df$job)[length(levels(df$job)) + 1] <- "missing"
# Then use the new level to indicate NA.
df$varname[is.na(df$job)] <- "missing"

df <- df %>% mutate(job = ifelse(is.na(job), yes = "Missing", no = job))

#marital
# Remove rows
df <- df[!is.na(df$marital), ]

sum(is.na(df))

summary(df)
df %>% mutate_if(is.character,as.factor) %>% summary()

df$varname <- NULL


```

```{r eval = F}
df %>% 
  mutate(edu_years = ifelse(is.na(edu_years), mean(df$edu_years), edu_years),
         job = ifelse(is.na(job), "missing", job)) %>% 
  filter(!is.na(housing),
         !is.na(marital)) %>% 
  summary()
```


The following code can be used to handle missing values.

```{r eval = F}

# Use one of the four choices below to deal with the NAs.
# Choose for each variable that has NAs.
# Replace varname with the actual variable name.

# Remove column
df$varname <- NULL

# Remove rows
df <- df[!is.na(df$varname), ]

# Convert missing values to "unknown" (works only for factor variables)
# First create a new level.
levels(df$varname)[length(levels(df$varname)) + 1] <- "unknown"
# Then use the new level to indicate NA.
df$varname[is.na(df$varname)] <- "unknown"

# Impute using the mean (works only for numeric variables)
df$varname[is.na(df$varname)] <- mean(df$varname, na.rm = TRUE)
```

# Task 4: Investigate correlations.

```{r} 
tmp <- dplyr::select(df, age, edu_years, CPI, CCI, irate, employment)
cor(tmp, use = "complete.obs")
```

# Task 5: Conduct a principal components analysis (PCA)

```{r}
# Perform Principal Components Analysis of the four variables below.
# The variables are standardized. Then the components are calculated.
tmp <- dplyr::select(df, CPI, CCI, irate, employment)
apply(tmp, 2, mean)
apply(tmp, 2, sd)
pca <- prcomp(tmp, scale = TRUE)
pca

# 	Create a bi-plot.
biplot(pca, cex = 0.8, xlabs = rep(".", nrow(tmp)))

# Consider the variance explained by the principal components.
# Use the output to decide how many principal components to use in the GLM models.
vars_pca <- apply(pca$x, 2, var)
vars_pca / sum(vars_pca)

plot(pca)
```

Calculate the principal components and add the first one to the data frame.

```{r}
pred <- as.data.frame(predict(pca, newdata = df[, c("CPI", "CCI", "irate", "employment")]))

df$PC1 <- pred$PC1
```

Split the data into training and testing. Check that the split looks reasonable.

```{r}
set.seed(1875)
train_ind <- createDataPartition(df$purchase, p = 0.7, list = FALSE)
data_train <- df[train_ind, ]
data_test <- df[-train_ind, ]
rm(train_ind)

print("Mean value of purchase on train and test data splits")
mean(data_train$purchase)
mean(data_test$purchase)

summary(df)
```

# Task 6: Create a generalized linear model (GLM).

>Don't panic.  You'll notice that the coefficients in this Rmd file are different that what is in the report.  You can expect this to happen to you and plan ahead.  It would a perfect if both matched 100% of the time, but in inevitably you will encounter these types of discrepencies.  Sometimes this is due to the random seed being different and different observations being used in the training and test sets.  It may also be due to differences in how the data was handled.  Perhaps the model was run on an earlier version of the data, and then I went back and made a change but did not update the values in the document.  **If you notice this, leave a comment such as this one in your Rmd file so that you'll still get full credit.**

```{r}
# Construct GLM using only age as an independent variable.

glm <- glm(purchase ~ age,
  data = data_train,
  family = binomial(link = "logit")
)

summary(glm)

# Evaluate GLM: construct ROC and calculate AUC
library(pROC)

glm_probs <- predict(glm, data_train, type = "response")
glm_probs_test <- predict(glm, data_test, type = "response")

roc <- roc(data_train$purchase, glm_probs)
par(pty = "s")
plot(roc)
pROC::auc(roc)

roc <- roc(data_test$purchase, glm_probs_test)
plot(roc)
pROC::auc(roc)

# Construct a GLM using a full set of independent variables and the first PC.

glm <- glm(purchase ~ age + job + marital + edu_years + housing + loan + phone + month + weekday + PC1,
  data = data_train,
  family = binomial(link = "logit")
)

summary(glm)

# Evaluate GLM: construct ROC and calculate AUC

glm_probs <- predict(glm, data_train, type = "response")
glm_probs_test <- predict(glm, data_test, type = "response")

roc <- roc(data_train$purchase, glm_probs)
par(pty = "s")
plot(roc)
pROC::auc(roc)

roc <- roc(data_test$purchase, glm_probs_test)
plot(roc)
pROC::auc(roc)
```

# Task 7: Select features using stepwise selection.

```{r}
# The following code executes the stepAIC procedure with backward selection and AIC. It first runs a new GLM adding the square of age.

glm2 <- glm(purchase ~ age + I(age^2) + job + marital + edu_years + housing + loan + phone + month + weekday + PC1,
  data = data_train,
  family = binomial(link = "logit")
)

library(MASS)


stepAIC(glm2,
  direction = "backward",
  k = 2
)

sum(is.na(data_train))

```

```{r}
# To evaluate the selected model, change the variable list to those selected by stepAIC.
glm.reduced <- glm(purchase ~ age + I(age^2) + edu_years + month + 
    weekday + PC1,
  data = data_train,
  family = binomial(link = "logit")
)

summary(glm.reduced)
```

# Task 8: Evaluate the model.

```{r}
# Evaluate GLM: construct ROC and calculate AUC

glm_probs <- predict(glm.reduced, data_train, type = "response")
glm_probs_test <- predict(glm.reduced, data_test, type = "response")

roc <- roc(data_train$purchase, glm_probs)
par(pty = "s")
plot(roc)
pROC::auc(roc)

roc <- roc(data_test$purchase, glm_probs_test)
plot(roc)
pROC::auc(roc)
```

```{r}
#Example: AUC = 0.5
#Randomly choose customers as 0 or 1 in the same proportion as the actual data
random_permutation <- sample(data_test$purchase, size = length(data_test$purchase), replace = F)
roc <- roc(data_test$purchase, random_permutation)
plot(roc)
pROC::auc(roc)
```

Example of AUC of 1

(This would not )

```{r}

glm_probs_test <- data_test$purchase

roc <- roc(data_test$purchase, glm_probs_test)
plot(roc)
pROC::auc(roc)
```


The following is an exxample of AUC of 0.  This would not need to be included in the actual Rmd. file but is included for illustration purposes.

```{r}
glm_probs_test <-  data_test$purchase

roc <- roc(data_test$purchase, glm_probs_test)
plot(roc)
pROC::auc(roc)
```

Check the base levels

```{r}
levels(as.factor(data_train$month))

levels(as.factor(data_train$weekday))
```

# Task 9: Investigate a shrinkage method.

```{r}
library(glmnet)
set.seed(42)


X.train <- model.matrix(purchase ~ age + I(age^2) + job + marital + edu_years + housing + loan + phone + month + weekday + PC1,
  data = data_train
)
X.test <- model.matrix(purchase ~ age + I(age^2) + job + marital + edu_years + housing + loan + phone + month + weekday + PC1,
  data = data_test
)

m <- cv.glmnet(
  x = X.train,
  y = data_train$purchase,
  family = "binomial",
  type.measure = "class",
  alpha = 0.5
) # alpha = 1 implies LASSO, alpha = 0 implies ridge, values between 0 and 1 imply elastic net
plot(m)
```

Use the cross-validation results to run the final elastic net regression model.

```{r}
# Fit the model
m.final <- glmnet(
  x = X.train,
  y = data_train$purchase,
  family = "binomial",
  lambda = m$lambda.min,
  alpha = 0.5
)

# List variables
m.final$beta

# Evaluate against train and test sets

# Predict on training data
enet.pred.train <- predict(m.final, X.train, type = "response")

roc <- roc(as.numeric(data_train$purchase), enet.pred.train[, 1])
par(pty = "s")
plot(roc)
pROC::auc(roc)

# Predict on test data
enet.pred.test <- predict(m.final, X.test, type = "response")

roc <- roc(as.numeric(data_test$purchase), enet.pred.test[, 1])
par(pty = "s")
plot(roc)
pROC::auc(roc)
```

# Task 10: Construct a decision tree.

```{r}

# Load the two needed libraries
library(rpart)
library(rpart.plot)

set.seed(1234)

formula <- "purchase ~ age + job + marital + edu_years + housing + loan + phone + month + weekday + CPI + CCI + irate"

tree1 <- rpart(formula,
  data = data_train, method = "class",
  control = rpart.control(minbucket = 5, cp = 0.0005, maxdepth = 7),
  parms = list(split = "gini")
)

rpart.plot(tree1, type = 0, digits = 4)

# Obtain predicted probabilities for train and for test.
pred.prob.tr <- predict(tree1, type = "prob")
pred.prob.te <- predict(tree1, type = "prob", newdata = data_test)

# Construct ROC and calculate AUC for the training data.
library(pROC)
print("Training ROC and AUC")
roc <- roc(data_train$purchase, pred.prob.tr[, "1"])
par(pty = "s")
plot(roc)

pROC::auc(roc)

# Do the same for test.
print("Test ROC and AUC")
roc2 <- roc(data_test$purchase, pred.prob.te[, "1"])
par(pty = "s")
plot(roc2)

pROC::auc(roc2)
```

# Task 11: Employ cost-complexity pruning to construct a smaller tree. 

```{r}

tree1$cptable # This code displays the complexity parameter table for tree1.
# Select the optimal pruning parameter from the table.
```

```{r}

# Replace XX in the code below with the selected complexity parameter.
tree2 <- prune(tree1, cp =  0.0061359346 , "CP")

# Show the pruned tree.
rpart.plot(tree2)

# Obtain predicted probabilities for train and for test.
pred.prune.prob.tr <- predict(tree2, type = "prob")
pred.prune.prob.te <- predict(tree2, type = "prob", newdata = data_test)

# Construct ROC and calculate AUC for the training data.
library(pROC)
print("Training ROC and AUC")
roc <- roc(data_train$purchase, pred.prune.prob.tr[, "1"])
par(pty = "s")
plot(roc)

pROC::auc(roc)

# Do the same for test.
print("Test ROC and AUC")
roc2 <- roc(data_test$purchase, pred.prune.prob.te[, "1"])
par(pty = "s")
plot(roc2)

pROC::auc(roc2)
```
