---
title: "Answers to Exercises - Data Manipulation with dplyr"
output:
  html_document:
    df_print: paged
---

```{r message = F}
library(ExamPAData)
library(tidyverse)
```


Run this code on your computer to answer these exercises.

The data `actuary_salaries` contains the salaries of actuaries collected from the DW Simpson survey.  Use this data to answer the exercises below.

```{r }
actuary_salaries %>% glimpse()
```

1.  How many industries are represented?
2.  The `salary_high` column is a character type when it should be numeric.  Change this column to numeric.
3.  What are the highest and lowest salaries for an actuary in Health with 5 exams passed?
4.  Create a new column called `salary_mid` which has the middle of the `salary_low` and `salary_high` columns.
5.  When grouping by industry, what is the highest `salary_mid`?  What about `salary_high`?  What is the lowest `salary_low`?
6.  There is a mistake when `salary_low == 11`.  Find and fix this mistake, and then rerun the code from the previous task.
7.  Create a new column, called `n_exams`, which is an integer.  Use 7 for ASA/ACAS and 10 for FSA/FCAS.  Use the code below as a starting point and fill in the `_` spaces
8. Create a column called `social_life`, which is equal to `n_exams`/`experience`.  What is the average (mean) `social_life` by industry?  Bonus question: what is wrong with using this as a statistical measure?


```{r eval = F, echo = T}
actuary_salaries <- actuary_salaries %>% 
  mutate(n_exams = case_when(exams == "FSA" ~ _,
                             exams == "ASA" ~ _,
                             exams == "FCAS" ~ _,
                             exams == "ACAS" ~ _,
                             TRUE ~ as.numeric(substr(exams,_,_)))) 
```

8. Create a column called `social_life`, which is equal to `n_exams`/`experience`.  What is the average (mean) `social_life` by industry?  Bonus question: what is wrong with using this as a statistical measure?


## Answers to exercises

1.  How many industries are represented?

```{r }
actuary_salaries %>% count(industry)
```

2.  The `salary_high` column is a character type when it should be numeric.  Change this column to numeric.

```{r}
#method 1
actuary_salaries <- actuary_salaries %>% 
  mutate(salary_high = as.numeric(salary_high))

#method 2
actuary_salaries <- actuary_salaries %>% 
  modify_at("salary_high", as.numeric)
```

3.  What are the highest and lowest salaries for an actuary in Health with 5 exams passed?

```{r }
actuary_salaries %>% 
  filter(industry == "Health", exams == 5) %>% 
  summarise(highest = max(salary_high),
            lowest = min(salary_low))
```


4.  Create a new column called `salary_mid` which has the middle of the `salary_low` and `salary_high` columns.

```{r  }
actuary_salaries <- actuary_salaries %>% 
  mutate(salary_mid = (salary_low + salary_high)/2)
```


5.  When grouping by industry, what is the highest `salary_mid`?  What about `salary_high`?  What is the lowest `salary_low`?

```{r  }
actuary_salaries %>% 
  group_by(industry) %>% 
  summarise(max_salary_mid = max(salary_mid),
            max_salary_high = max(salary_high),
            low_salary_low = min(salary_low))
```

6.  There is a mistake when `salary_low == 11`.  Find and fix this mistake, and then rerun the code from the previous task.

```{r }
actuary_salaries <- actuary_salaries %>% 
  mutate(salary_low = ifelse(salary_low == 11, 
                             yes = 114, no = salary_low),
         salary_high = ifelse(salary == "114 - 166", 
                              yes = 166, no = salary_high))

#the minimum salary low is now 48
actuary_salaries %>% 
  group_by(industry) %>% 
  summarise(max_salary_mid = max(salary_mid),
            max_salary_high = max(salary_high),
            low_salary_low = min(salary_low))
```

7.  Create a new column, called `n_exams`, which is an integer.  Use 7 for ASA/ACAS and 10 for FSA/FCAS.

Use the code below as a starting point and fill in the `_` spaces

```{r eval = F, echo = T}
actuary_salaries <- actuary_salaries %>% 
  mutate(n_exams = case_when(exams == "FSA" ~ _,
                             exams == "ASA" ~ _,
                             exams == "FCAS" ~ _,
                             exams == "ACAS" ~ _,
                             TRUE ~ as.numeric(substr(exams,_,_)))) 
```


```{r  }
actuary_salaries <- actuary_salaries %>% 
  mutate(n_exams = case_when(exams == "FSA" ~ 10,
                             exams == "ASA" ~ 7,
                             exams == "FCAS" ~ 10,
                             exams == "ACAS" ~ 7,
                             TRUE ~ as.numeric(substr(exams,1,2)))) 

actuary_salaries %>% count(n_exams)
```

8. Create a column called `social_life`, which is equal to `n_exams`/`experience`.  What is the average (mean) `social_life` by industry?  Bonus question: what is wrong with using this as a statistical measure?

```{r}
actuary_salaries %>% 
  mutate(social_life = n_exams/experience) %>% 
  group_by(industry) %>% 
  summarise(avg_social_life = mean(social_life))
```

This is not REALLY an average as the number of people, or number of actuaries, are not taken into consideration.  To correctly take the average, we would need use a weighted average like sum(number_actuaries*social_life)/sum(number_actuaries)

