---
title: "GLM Short-Answer Practice Questions"
output:
  html_document:
    df_print: paged
---

The following code is needed just for setting up your practice questions.  No changes are needed.

```{r setup, include=F, eval=T}
library(ExamPAData)
library(caret)
library(plyr)
library(dplyr)

## Patient Length of Stay ##

data.all <- patient_length_of_stay

data.all$age <- as.factor(data.all$age) # Replace VARIABLE twice
data.all$admit_type_id <- as.factor(data.all$admit_type_id) # Replace VARIABLE twice

data.all <- data.all %>% filter(gender != "Unknown/Invalid:")

data.all <- data.all %>% filter(gender %in% c("Male", "Female"))

#gender_missing <- data.all$gender == "Unknown/Invalid:"
#data.all <- data.all[-gender_missing,]

data.all <- data.all %>% mutate(race = ifelse(race == "?", "Other", race))

#data.all$race <- ifelse(data.all$race == "?", "Other", data.all$race)

data.all$weight <- NULL # Replace VARIABLE

data.all <- data.all %>% mutate(race = ifelse(race == "Asian", "Other", race))
data.all$race <- ifelse(data.all$race=="Asian", "Other", data.all$race)

library(forcats)
data.all <- data.all <- data.all %>% mutate_if(is.factor, fct_infreq)

data.all <- data.all <- data.all %>% mutate_if(is.character, fct_infreq)

# Select only the variables used for PCA. Do not change this list.
data.pca <- data.all[, c("num_procs", "num_meds", "num_ip", "num_diags")]

# Run PCA on the numeric variables. Variables are centered and scaled.
pca <- prcomp(data.pca, center = TRUE, scale. = TRUE)

# Center and scale the variables
data.pca.std <- as.data.frame(scale(data.pca))
head(data.pca.std)

# Add the first principal component to the data frame.
data.all$PC1 <- pca$x[, 1]

set.seed(100)
train_ind <- createDataPartition(data.all$days, p = 0.7, list = FALSE)
data.train.los <- data.all[train_ind, ]
data.test.los <- data.all[-train_ind, ]

## Health Costs

theme_set(theme_bw())
set.seed(1234)

#set factor levels to those with the most observations
health_insurance <- health_insurance %>% 
  mutate(sex = fct_infreq(sex),
         smoker = fct_infreq(smoker),
         region = fct_infreq(region),
         )

#your code here
health_insurance <- health_insurance%>% 
  mutate(age_bucket = case_when(
    age < 24 ~ 1,
    age <= 36 ~ 2,
    age <= 50 ~ 3,
    age > 50 ~ 4
  ) %>% as.character() %>% fct_infreq(),
  age_household_ratio = (1 + children)/age,
  log_household_size = log(1 + children)
  ) 

## Customer Value ##

library(kableExtra)
theme_set(theme_minimal())

df <- customer_value %>% 
  mutate(value_flag = ifelse(value_flag=="High",1,0))

df <- df %>% dplyr::mutate(marital_status = ifelse(marital_status == "Married-AF-spouse", 
                                            yes = "Married-civ-spouse",
                                            no = marital_status))

df$marital_status <- as.factor(df$marital_status)

simple_levels <- c("Divorced", "Married-spouse", "Married-spouse", "Never-married", "Separated", "Widowed")

#Combine the two marital status levels
var.levels <- levels(df$marital_status)
df$marital_status <- mapvalues(df$marital_status,var.levels, simple_levels) 

#Change base level
table <- as.data.frame(table(df$marital_status))
max <- which.max(table[,2])#base level is one with most observations
level.name <- as.character(table[max,1])
df$marital_status <- relevel(df$marital_status, ref = level.name)

df <- df %>% mutate_if(is.character, fct_infreq)

train_ind <- createDataPartition(df$value_flag, p = 0.7, list = FALSE)
data.train <- df[train_ind, ]
data.test <- df[-train_ind, ]

library(rpart)
library(rpart.plot)

# This code takes a continuous variable and creates a binned factor variable. The code applies it directly to the capital gain variable as an example. right = FALSE means that the left number is included and the right number excluded. So, in this case, the first bin runs from 0 to 1000 and includes 0 and excludes 1000. Note that the code creates a new variable, so the original variable is retained.
df$cap_gain_cut <- cut(df$cap_gain, breaks = c(0, 5000, 7000, Inf), right = FALSE, labels = c("lowcg", "mediumcg", "highcg"))

#df <- df %>% select(-cap_gain)
train_ind <- createDataPartition(df$value_flag, p = 0.7, list = FALSE)
data.train <- df[train_ind, ]
data.test <- df[-train_ind, ]
print("TRAIN")
mean(data.train$value_flag)
print("TEST")
mean(data.test$value_flag)
rm(train_ind)
```

# Question 1

Fit a GLM with all original variables included but excluding the PCA variable. Do not change the Poisson distribution or log link function.

```{r echo=T}
# Fit the model
glm1 <- glm(days ~ . - PC1,
  data = data.train.los,
  family = poisson(link = "log") # Do not change.
)

summary(glm1)

# Predict on training data
glm1.pred.train <- predict(glm1, data.train.los, type = "response")

# Calculate the Pearson goodness-of-fit statistic on training data
sum((data.train.los$days - glm1.pred.train)^2 / glm1.pred.train) / nrow(data.train.los)

# Predict on test data
glm1.pred.test <- predict(glm1, data.test.los, type = "response")

# Calculate the Pearson goodness-of-fit statistic on test data
sum((data.test.los$days - glm1.pred.test)^2 / glm1.pred.test) / nrow(data.test.los)
```

Fit a GLM with the PCA variable created in Task 5 included (and without the numeric variables used to produce the PCA variable). Do not change the Poisson distribution or log link function.

```{r echo=T}
# Fit the model
glm2 <- glm(days ~ . - num_procs - num_meds - num_ip - num_diags,
  data = data.train.los,
  family = poisson(link = "log") # Do not change.
)

summary(glm2)

# Predict on training data
glm2.pred.train <- predict(glm2, data.train.los, type = "response")

# Calculate the Pearson goodness-of-fit statistic on training data
sum((data.train.los$days - glm2.pred.train)^2 / glm2.pred.train) / nrow(data.train.los)

# Predict on test data
glm2.pred.test <- predict(glm2, data.test.los, type = "response")

# Calculate the Pearson goodness-of-fit statistic on test data
sum((data.test.los$days - glm2.pred.test)^2 / glm2.pred.test) / nrow(data.test.los)
```


# Question 2

```{r message = F, echo=T}
#create a train/test split
index <- createDataPartition(y = health_insurance$charges, list = F, p = 0.8) %>% as.numeric()
train <-  health_insurance %>% slice(index)
test <- health_insurance %>% slice(-index)

nrow(train) + nrow(test) == nrow(health_insurance)
nrow(train)/nrow(health_insurance)
```


```{r echo=T}
#your assistant has provided you with these functions
#no changes need to be made
mae <- function(y, y_hat){
  mean(abs(y - y_hat))
}

fit_glm <- function(family_link = gaussian(link = "identity")){
  model = glm(formula = charges ~ age + sex + bmi + smoker + region + age_bucket + age_household_ratio + log_household_size + bmi*smoker, 
            family = family_link,
            data = train)
  pred = predict(model, test, type = "response")
  paste0(
    "family:", model$family$family, ", ",
    "link:", model$family$link, ", ",
    "test mae:", round(mae(pred, test$charges),0))
}
```

For example, to fit a GLM with a Gaussian response and a log link, use

```{r echo=T}
fit_glm(gaussian(link = "log"))
```

```{r echo=T}
fit_glm(gaussian(link = "identity"))
fit_glm(gaussian(link = "log"))
fit_glm(gaussian(link = "sqrt"))
fit_glm(gaussian(link = "1/mu^2"))
fit_glm(gaussian(link = "inverse"))
```

```{r echo=T}
fit_glm(Gamma(link = "identity"))
fit_glm(Gamma(link = "log"))
fit_glm(Gamma(link = "sqrt"))
fit_glm(Gamma(link = "1/mu^2"))
fit_glm(Gamma(link = "inverse"))
```

```{r echo=T}
fit_glm(inverse.gaussian(link = "identity"))
fit_glm(inverse.gaussian(link = "log"))
fit_glm(inverse.gaussian(link = "sqrt"))
#fit_glm(inverse.gaussian(link = "1/mu^2"))
#fit_glm(inverse.gaussian(link = "inverse"))
```


To find the possible families, see `?family`.

The possible link functions are `logit`, `probit`, `cauchit`, `cloglog`, `identity`, `log`, `sqrt`, `1/mu^2`, and `inverse`.

# Question 3

No code provided (short-answer only).

# Question 4

```{r echo=T}
names(data.train)
```

```{r echo=T}
glm <- glm(value_flag ~ age + education_num + marital_status + occupation + cap_gain_cut + hours_per_week + score + marital_status*cap_gain_cut,
  data = data.train,
  family = binomial(link = "logit")
)


# Replace VARIABLE LIST with selected variables, DISTRIBUTION with the distribution choice, and LINK with the link choice

summary(glm)

glm.probs <- predict(glm, data.train, type = "response")
glm.pred <- glm.probs > 0.5
table(glm.pred, data.train$value_flag)

glm.probs.test <- predict(glm, data.test, type = "response")
glm.pred.test <- glm.probs.test > 0.5
table(glm.pred.test, data.test$value_flag)

# Construct ROC and calculate AUC for the training data.
library(pROC)
roc <- roc(data.train$value_flag, glm.probs)
par(pty = "s")
plot(roc)

pROC::auc(roc)

# Construct ROC and calculate AUC for the test data.

roc <- roc(data.test$value_flag, glm.probs.test)

plot(roc)

pROC::auc(roc)
```

The following code executes the stepAIC procedure allowing for choices of AIC/BIC and forward/backward.

```{r echo=T}
# This code runs the stepAIC procdure. It is set up to use the GLM model previously fit along with forward #selection and BIC. That does not imply these are the best choices.

# Note: When using BIC, decisions about adding or removing variables are made using that criterion. However, the #AIC value presented at the end when the final model is run uses the standard AIC penalty of k = 2.

library(MASS)
# If using forward selection it is necessary to fit a model with no predictors to use as the start.
glm.none <- glm(value_flag ~ 1, data = data.train, family = binomial(link = "logit"))

#forward

stepAIC(glm.none,
  direction = "forward",
  k = 2, #log(nrow(data.train)),
  scope = list(upper = glm, lower = glm.none) # set k = 2 for AIC
)

# backward
stepAIC(glm,
  direction = "backward",
  k = 2,
  scope = list(upper = glm, lower = glm.none) # set k = 2 for AIC
)
```

