---
title: "Practice Exam - Health Care Costs"
output: word_document
editor_options:
  chunk_output_type: inline
---

### Task 1 - Write a high-level summary of the data 

```{r message = F}
#set up
library(tidyverse)
library(ExamPAData)
library(gbm)
library(caret)

theme_set(theme_bw())
set.seed(1234)

#set factor levels to those with the most observations
health_insurance <- health_insurance %>% 
  mutate(sex = fct_infreq(sex),
         smoker = fct_infreq(smoker),
         region = fct_infreq(region),
         )
```

```{r}
#start here

```


### Task 2 - Explore the target variables CHARGES and its relationship with SMOKER and BMI 

```{r}
health_insurance %>% 
  ggplot(aes(charges)) +
  geom_histogram()

health_insurance$charges %>% summary()
```

```{r}
health_insurance %>% 
  ggplot(aes(charges, fill = smoker)) +
  geom_histogram()
```


```{r}
health_insurance %>% 
  ggplot(aes(charges)) + 
  geom_histogram() + 
  facet_wrap(vars(smoker))
```

```{r}
health_insurance %>% 
  ggplot(aes(charges, bmi)) + 
  geom_point()
```

```{r}
health_insurance %>% 
  ggplot(aes(charges, bmi, color = smoker)) + 
  geom_point()
```


### Task 3 - Engineer additional features

```{r}
#your code here
```

### Task 4 - Hierarchical Clustering

```{r fig.width=12}
age <- health_insurance$age[1:50]
bmi <- health_insurance$bmi[1:50]

#your code here
#apply scaling
#(no code is provided)

#put age and bmi into a matrix
x <- as.matrix(tibble(age = age, bmi = bmi))

#perform hiearchical clustering
hc_average = hclust(dist(x), method = "average")

#plot the dedgrogram
plot(hc_average, main = "Average Linkage", xlab = "", sub = "", cex = .9)
```

Select which clustering method to use as well as the number of clusters.

```{r}
#make a selection
selected_height = 0.6
```



```{r}
cluster_data <- tibble(age = age, bmi = bmi) %>% 
  mutate(cluster = as.factor(cutree(hc_average,h= selected_height))) 

cluster_data %>% 
  ggplot(aes(age, bmi, color = cluster)) + 
  geom_point()
```

### Task 5 - Select an Interaction

```{r}
#example of a formula using all variables + an interaction between age and sex
glm_formula = charges ~ age + sex + bmi + smoker + region + charges + age_bucket + age_household_ratio + log_household_size + age*sex
```


### Task 6 - Fit two GLMs 

```{r message = F}
#create a train/test split
index <- createDataPartition(y = health_insurance$charges, list = F, p = 0.8) %>% as.numeric()
train <-  health_insurance %>% slice(index)
test <- health_insurance %>% slice(-index)

nrow(train) + nrow(test) == nrow(health_insurance)
nrow(train)/nrow(health_insurance)
```


```{r}
#your assistant has provided you with these functions
#no changes need to be made
mae <- function(y, y_hat){
  mean(abs(y - y_hat))
}

fit_glm <- function(family_link = gaussian(link = "identity")){
  model = glm(formula = glm_formula, 
            family = family_link,
            data = train)
  pred = predict(model, test, type = "response")
  paste0(
    "family:", model$family$family, ", ",
    "link:", model$family$link, ", ",
    "mae:", round(mae(pred, test$charges),0))
}
```

For example, to fit a GLM with a Gaussian response and a log link, use

```{r}
fit_glm(gaussian(link = "log"))
```

To find the possible families, see `?family`.

The possible link functions are `logit`, `probit`, `cauchit`, `cloglog`, `identity`, `log`, `sqrt`, `1/mu^2`, and `inverse`.

### Task 7 - Compare the two models


### Task 8 - Examine the residual plots


```{r fig.width=12, fig.height=4}
#change the family and link to the model that you choose
model = glm(formula = glm_formula, 
            family = gaussian(link = "log"),
            data = train)

pred = predict(model, test, type = "response")
mae(pred, test$charges)
par(mfrow = c(1,3))
plot(model, which = c(1,2,3))
summary(model)
```

### Task 9 - Fit a GBM

Your assistant has already run cross-validation to find the best set of hyperparameters.  This is

 n.trees interaction.depth shrinkage n.minobsinnode
 1000                 3      0.01             30

```{r}
#no changes are needed to this code

#look at many different parameter combinations
tunegrid <- expand.grid(
    interaction.depth = c(3),
    n.trees = c(1000), 
    shrinkage = c(0.01),
    n.minobsinnode = c(30)
    )

#use the mae as the evaluation metric
mae_summary <- function (data,
                        lev = NULL,
                        model = NULL) {
      out <- mae(data$obs, data$pred)  
      names(out) <- "mae"
      out
}

#use 3-fold cross validation
control <- trainControl(
  summaryFunction = mae_summary,
  method='repeatedcv', 
  number=3,  
  p = 0.8)

#fit a GBM using MAE as the error metric
gbm <- train(charges ~ .,
            data = train,
            method='gbm', 
            tuneGrid=tunegrid, 
            metric = "mae",
            maximize = FALSE,
            trControl=control,
            verbose = FALSE
            )

```

### Task 10 - Validate the GBM

```{r}
#get the mae on the test set
pred = predict(gbm, test)
mae(pred, test$charges)
```

### Task 11 - Interpret the GBM

```{r fig.width=5, fig.height=4}
par(las = 1, cex = 0.5)
summary(gbm)
```

```{r}
pdp::partial(gbm, pred.var = "age", grid.resolution = 20, plot = T)
pdp::partial(gbm, pred.var = "bmi", grid.resolution = 20, plot = T)

pdp::partial(gbm, pred.var = c("smoker", "age"), grid.resolution = 20, plot = T)
pdp::partial(gbm, pred.var = c("smoker", "bmi"), grid.resolution = 20, plot = T)
```

### Task 12 - Executive Summary

Write in the Word template.