---
title: "Health Costs Solution"
subtitle: "ExamPA.net Practice Exam"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    theme: cerulean
    toc: true
    code_folding: show
    toc_float: true
---

For your convenience, this solution document has been formatted to be easy to read.  When taking this exam, candidates are not expected to make their `.Rmd` documents look this way.  The `make_table` function is included only to improve your reading experience.  The SOA's instructions from the Project Statement say only that

>At a minimum you must submit your completed report template and an Rmd file that supports your work.  Graders expect that your Rmd code can be run from beginning to end.  The code snippets provided should either be commented out or adapted for execution.  Ensure that it is clear where in the code each of the tasks is addressed.  Documentation of your thought process and conclusion for each task should be completely contained within your Word report.  The Rmd code should be clear, may contain commentary, and support your work.

### Task 1 - Write a high-level summary of the data 

```{r message = F}
#Set up
#No changes needed to this code chunk

#set up
library(tidyverse)
library(ExamPAData)
library(gbm)
library(caret)
library(kableExtra)

#function to create html tables
make_table <- function(df, caption = ""){
  df %>% 
    kable(caption = caption) %>% 
    kable_styling(bootstrap_options = "striped", full_width = F)}

#global chunk options
knitr::opts_chunk$set(warning = F, message = F, echo = T, fig.width = 14, fig.height = 8, dev.args = list(type = "cairo"))

theme_set(theme_bw())
set.seed(1234)

#set factor levels to those with the most observations
health_insurance <- health_insurance %>% 
  mutate(sex = fct_infreq(sex),
         smoker = fct_infreq(smoker),
         region = fct_infreq(region),
         )
```

```{r}
#start here
health_insurance %>% glimpse()

#Can also use
#summary(health_insurance)
#str(health_insurance)
```


### Task 2 - Explore the target variables CHARGES and its relationship with SMOKER and BMI 

```{r}
health_insurance %>% 
  ggplot(aes(charges)) +
  geom_histogram()

health_insurance$charges %>% summary()

health_insurance %>% ggplot(aes(bmi)) + geom_histogram()
```

```{r}
health_insurance %>% 
  ggplot(aes(charges, fill = smoker)) +
  geom_histogram()

health_insurance %>% 
  group_by(smoker) %>% 
  summarise(mean = mean(charges),
            median = median(charges)) %>% 
  make_table()
```


```{r}
health_insurance %>% 
  ggplot(aes(charges)) + 
  geom_histogram() + 
  facet_wrap(vars(smoker))
```

```{r}
health_insurance %>% 
  ggplot(aes(charges, bmi)) + 
  geom_point()
```

```{r}
health_insurance %>% 
  ggplot(aes(charges, bmi, color = smoker)) + 
  geom_point()
```


### Task 3 - Engineer additional features

```{r}
#your code here
health_insurance <- health_insurance%>% 
  mutate(age_bucket = case_when(
    age < 24 ~ 1,
    age <= 36 ~ 2,
    age <= 50 ~ 3,
    age > 50 ~ 4
  ) %>% as.character() %>% fct_infreq(),
  age_household_ratio = (1 + children)/age,
  log_household_size = log(1 + children)
  ) %>% 
  select(-children) 

health_insurance %>% select(age_bucket, age_household_ratio, log_household_size) %>% summary()

health_insurance %>% count(age_bucket) %>% make_table
```

```{r eval = F}
#Base R solution

#See documentation of ?cut
#health_insurance$age_bucket2 <- cut(x = health_insurance$age, breaks = c(0,23, 36, 50, 500))

#table(health_insurance$age_bucket2)
```


### Task 4 - Hierarchical Clustering

```{r}
health_insurance %>% ggplot(aes(age, bmi)) + geom_point()
health_insurance %>% summary()

#euclidean distance example
x0 <- 40
y0 <- 40

x1 <- 60
y1 <- 50

sqrt(sum())

sqrt((40 - 60)^2 + (40-50)^2)


```


```{r fig.width=12}
age <- health_insurance$age[1:50]
bmi <- health_insurance$bmi[1:50]

#your code here
#apply scaling
#(no code is provided)
age <- (age - mean(age))/sd(age)
bmi <- (bmi - mean(bmi))/sd(bmi)

summary(age)
summary(bmi)

#put age and bmi into a matrix
x <- as.matrix(tibble(age = age, bmi = bmi))

#perform hiearchical clustering
hc_average = hclust(dist(x), method = "average")

#plot the dedgrogram
plot(hc_average, main = "Average Linkage", xlab = "", sub = "", cex = .9)
```

Select which clustering method to use as well as the number of clusters.

```{r}
#make a selection
selected_height = 1
```



```{r}
cluster_data <- tibble(age = age, bmi = bmi) %>% 
  mutate(cluster = as.factor(cutree(hc_average,h= selected_height))) 

cluster_data %>% 
  ggplot(aes(age, bmi, color = cluster)) + 
  geom_point()
```

### Task 5 - Select an Interaction

```{r}
#example of a formula using all variables + an interaction between age and sex
glm_formula = charges ~ age + sex + bmi + smoker + region + age_bucket + age_household_ratio + log_household_size + age*sex

glm_formula = charges ~ age + sex + bmi + smoker + region  + age_bucket + age_household_ratio + log_household_size + bmi*smoker
```


### Task 6 - Fit two GLMs 

```{r message = F}
#create a train/test split
index <- createDataPartition(y = health_insurance$charges, list = F, p = 0.8) %>% as.numeric()
train <-  health_insurance[index,]
test <- health_insurance[-index,]

nrow(train) + nrow(test) == nrow(health_insurance)
nrow(train)/nrow(health_insurance)
```


```{r}
#your assistant has provided you with these functions
#no changes need to be made
mae <- function(y, y_hat){
  mean(abs(y - y_hat))
}

fit_glm <- function(family_link = gaussian(link = "identity")){
  model = glm(formula = glm_formula, 
            family = family_link,
            data = train)
  pred = predict(model, test, type = "response")
  paste0(
    "family:", model$family$family, ", ",
    "link:", model$family$link, ", ",
    "test mae:", round(mae(pred, test$charges),0))
}
```

For example, to fit a GLM with a Gaussian response and a log link, use

```{r}
fit_glm(gaussian(link = "log"))
```

```{r}
fit_glm(gaussian(link = "identity"))
fit_glm(gaussian(link = "log"))
fit_glm(gaussian(link = "sqrt"))
fit_glm(gaussian(link = "1/mu^2"))
fit_glm(gaussian(link = "inverse"))
```

```{r}
fit_glm(Gamma(link = "identity"))
fit_glm(Gamma(link = "log"))
fit_glm(Gamma(link = "sqrt"))
fit_glm(Gamma(link = "1/mu^2"))
fit_glm(Gamma(link = "inverse"))
```

```{r}
fit_glm(inverse.gaussian(link = "identity"))
fit_glm(inverse.gaussian(link = "log"))
fit_glm(inverse.gaussian(link = "sqrt"))
#fit_glm(inverse.gaussian(link = "1/mu^2"))
#fit_glm(inverse.gaussian(link = "inverse"))
```


To find the possible families, see `?family`.

The possible link functions are `logit`, `probit`, `cauchit`, `cloglog`, `identity`, `log`, `sqrt`, `1/mu^2`, and `inverse`.

### Task 7 - Compare the two models  


### Task 8 - Examine the residual plots


```{r fig.width=12, fig.height=4}
#change the family and link to the model that you choose
model = glm(formula = glm_formula, 
            family = Gamma(link = "identity"),
            data = train)

pred = predict(model, test, type = "response")
mae(pred, test$charges)
par(mfrow = c(1,3))
plot(model, which = c(1,2,3))
summary(model)
```

### Task 9 - Fit a GBM

Your assistant has already run cross-validation to find the best set of hyperparameters.  This is

 n.trees interaction.depth shrinkage n.minobsinnode
 1000                 3      0.01             30

```{r}
#no changes are needed to this code

#look at many different parameter combinations
tunegrid <- expand.grid(
    interaction.depth = c(3),
    n.trees = c(1000), 
    shrinkage = c(0.01),
    n.minobsinnode = c(30)
    )

#use the mae as the evaluation metric
mae_summary <- function (data,
                        lev = NULL,
                        model = NULL) {
      out <- mae(data$obs, data$pred)  
      names(out) <- "mae"
      out
}

#use 3-fold cross validation
control <- trainControl(
  summaryFunction = mae_summary,
  method='repeatedcv', 
  number=3,  
  p = 0.8)

#fit a GBM using MAE as the error metric
gbm <- train(charges ~ .,
            data = train,
            method='gbm', 
            tuneGrid=tunegrid, 
            metric = "mae",
            maximize = FALSE,
            trControl=control,
            verbose = FALSE
            )

```

### Task 10 - Validate the GBM

```{r}
#get the mae on the test set
pred = predict(gbm, test)
mae(pred, test$charges)
```

### Task 11 - Interpret the GBM

```{r fig.width=5, fig.height=4}
par(las = 1, cex = 0.5)
summary(gbm)
```

```{r}
pdp::partial(gbm, pred.var = "age", grid.resolution = 20, plot = T)
pdp::partial(gbm, pred.var = "bmi", grid.resolution = 20, plot = T)

pdp::partial(gbm, pred.var = c("smoker", "age"), grid.resolution = 20, plot = T)
pdp::partial(gbm, pred.var = c("smoker", "bmi"), grid.resolution = 20, plot = T)
```

### Task 12 - Executive Summary

Write in the Word template.