---
title: "Hospital Readmissions - solution"
subtitle: "Futuro Insight Model"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    theme: cerulean
    toc: true
    code_folding: show
    toc_float: true
---

This chunk performs binarization. Note that it is set to fullRank = FALSE. This creates binarized variables for each factor level. If set to TRUE it will leave one out. Note the new variables are placed in a new dataframe. It can attached to an existing dataframe via old.df <- cbind(old.df, binarized_vars)

```{r echo = F, warning = F, message = F}
#Set up 
#No changes needed
library(ExamPAData)
library(ggplot2)
library(readr)
library(tidyr)
library(broom)
library(forcats)
library(caret)
library(gridExtra)
library(rpart)
library(rpart.plot)
library(caret)
library(kableExtra)
library(dplyr)
theme_set(theme_minimal())
#function to create html tables
make_table <- function(df, caption = ""){
  df %>% 
    kable(caption = caption) %>% 
    kable_styling(bootstrap_options = "striped", full_width = F)}
```


```{r eval = F}
factor_names <- c("ER","Age") #insert the column names of the variables to be binarized
factor_vars <- readmission[,factor_names]
for (var in factor_names) {
  factor_vars[, var] <- as.character(factor_vars[, var])
}

binarizer <- caret::dummyVars(paste("~", paste(factor_names, collapse = "+")) , data = factor_vars, fullRank = TRUE)
binarized_vars <- data.frame(predict(binarizer, newdata = factor_vars))
head(binarized_vars)
```

This chunk creates training and testing sets.

```{r eval = F}
#Create train and test sets
set.seed(4321)
partition <- createDataPartition(readmission[,1], list = FALSE, p = .75) #The partition will stratify using variable 1 from the dataframe
train <- readmission[partition, ]
test <- readmission[-partition, ]

print("TRAIN")
mean(train$Readmission.Status)

print("TEST")
mean(test$Readmission.Status)
```

The following chunk provides code that can be used to combine factor levels. It also relevels in case the new level has the highest frequency. 

```{r eval = F}
#USED fct-infreq method instead

#This example combines levels other than White of Race into a new level called NonWhite.
#Execute the function levels(readmission$Race) to identify the levels. Be sure the variable is a factor variable before doing this. This code assumes the variable has previously been releveled so that "White" is the first level.

#readmission2<-readmission #The results are in a new data frame called readmission2. This is done so that the results can be checked without losing the original data frame. When done, consider executing readmission <- readmission2

#Warning: If you run this code and import the plyr library, it will cause errors later on.
#To fix these errors, there are two options
#1. Restart R and RStudio and load tidyverse AFTER plyr
#2. Tell R to use the dplyr library for functions that share the same name in both libraries: dplyr::summarise, dplyr::group_by, etc.
#library(plyr)
#var <- "Race"
#var.levels <- levels(readmission2[,var])
#readmission2[,var] <- mapvalues(readmission2[,var],var.levels,c("White","NonWhite","NonWhite","NonWhite"))
#Relevel
#table <- as.data.frame(table(readmission2[,var]))
#  max <- which.max(table[,2])
#  level.name <- as.character(table[max,1])
#  readmission2[,var] <- relevel(readmission2[,var], ref = level.name)

#table(readmission2[,var])
```


This chunk reads in the data, relevels factors, and prints a summary.

```{r}
#Set factor levels to those with the most observations
readmission <- readmission %>% 
  mutate_if(is.character, fct_infreq)

summary(readmission) 
glimpse(readmission)
```


## Task 1 -  Perform univariate exploration of the four non-factor variables (6 points)

Code is provided to create a histogram for one of the variables.

```{r}
library(ggplot2)
readmission %>% ggplot(aes(ER)) + geom_histogram(bins = 9) #Set the number of bins equal to 9 because there are 9 unique values
readmission %>% ggplot(aes(LOS)) + geom_histogram(bins = 36) #Set the number of bins equal to 36 because there are 36 unique values
readmission %>% ggplot(aes(log(LOS))) + geom_histogram()

readmission %>% ggplot(aes(Age)) + geom_histogram()
readmission %>% ggplot(aes(HCC.Riskscore)) + geom_histogram()

readmission <- readmission %>% mutate(log_LOS = log(LOS),
                                      log_riskscore = log(HCC.Riskscore)) %>% 
  select(-LOS, -HCC.Riskscore)
```


## Task 2 - Examine relationships between DRG.Class and DRG.Complication (5 points)

Code is provided to create a tabular view of the two variables.

```{r}
readmission %>% 
  count(DRG.Class, DRG.Complication) %>% 
  spread(DRG.Class,n) %>% 
  make_table()

#Using base R
table(readmission$DRG.Class,readmission$DRG.Complication) 

#Using base R
#rows_to_drop <- (readmission$DRG.Complication  == "MedicalMCC.CC") & (readmission$DRG.Class == "SURG")
#readmission_temp <- readmission[!rows_to_drop,]

#removes the 6 patients 
readmission <- readmission %>% 
  filter(!(DRG.Complication == "MedicalMCC.CC" & DRG.Class == "SURG"))

readmission <- readmission %>% 
  mutate(DRG.Class = case_when(
    DRG.Complication == "MedicalMCC.CC" & DRG.Class == "MED" ~ "MED",
    DRG.Complication == "SurgMCC.CC" & DRG.Class == "SURG" ~ "SURG",
    DRG.Complication == "MedicalNoC" & DRG.Class == "MED" ~ "MED",
    DRG.Complication == "SurgNoC" & DRG.Class == "SURG" ~ "SURG",
    TRUE ~ "OTHER"
  )) 

#Using Base R,
#readmission_temp$DRG.Class <- case_when(
#    readmission$DRG.Complication == "MedicalMCC.CC" & readmission$DRG.Class == "MED" ~ "MED",
#    readmission$DRG.Complication == "SurgMCC.CC" & readmission$DRG.Class == "SURG" ~ "SURG",
#    readmission$DRG.Complication == "MedicalNoC" & readmission$DRG.Class == "MED" ~ "MED",
#    readmission$DRG.Complication == "SurgNoC" & readmission$DRG.Class == "SURG" ~ "SURG",
#    TRUE ~ "OTHER"
#  ) 


readmission %>% count(DRG.Class) %>% make_table()
#table(readmission_temp$DRG.Class)

#relevel the new variable
readmission$DRG.Class <- fct_infreq(readmission$DRG.Class)
```

## Task 3 - Use observations from cluster analysis to consider a new feature (9 points)

Code is provided to perform cluster analysis for from 1 to 12 clusters, construct an elbow plot and create a new variable based on a selected number of clusters. That variable will need to be retained for potentially being added tot he dataframe.

```{r}
nstart.val <- 20
cluster_vars <- readmission[c('log_LOS','Age')]
for(i in 1:ncol(cluster_vars)){
  cluster_vars[,i] <- scale(cluster_vars[,i])
}
km1 <- kmeans(cluster_vars,centers=1,nstart=nstart.val)
km2 <- kmeans(cluster_vars,centers=2,nstart=nstart.val)
km3 <- kmeans(cluster_vars,centers=3,nstart=nstart.val)
km4 <- kmeans(cluster_vars,centers=4,nstart=nstart.val)
km5 <- kmeans(cluster_vars,centers=5,nstart=nstart.val)
km6 <- kmeans(cluster_vars,centers=6,nstart=nstart.val)
km7 <- kmeans(cluster_vars,centers=7,nstart=nstart.val)
km8 <- kmeans(cluster_vars,centers=8,nstart=nstart.val)
km9 <- kmeans(cluster_vars,centers=9,nstart=nstart.val)
km10 <- kmeans(cluster_vars,centers=10,nstart=nstart.val)
km11 <- kmeans(cluster_vars,centers=11,nstart=nstart.val)
km12 <- kmeans(cluster_vars,centers=12,nstart=nstart.val)

var.exp <- data.frame(k = c(1:12),
                      bss_tss = c(km1$betweenss/km1$totss,
                                  km2$betweenss/km2$totss,
                                  km3$betweenss/km3$totss,
                                  km4$betweenss/km4$totss,
                                  km5$betweenss/km5$totss,
                                  km6$betweenss/km6$totss,
                                  km7$betweenss/km7$totss,
                                  km8$betweenss/km8$totss,
                                  km9$betweenss/km9$totss,
                                  km10$betweenss/km10$totss,
                                  km11$betweenss/km11$totss,
                                  km12$betweenss/km12$totss))

ggplot(var.exp,aes(x=k,y=bss_tss))+geom_point()

LOS_Age_Clust <- as.factor(km4$cluster) #This creates a new variable based on having 8 clusters.
cluster_vars$LOS_Age_Clust <- LOS_Age_Clust
ggplot(data = cluster_vars, aes(x = Age, y = log_LOS, col = LOS_Age_Clust)) + geom_point() + theme(axis.text = element_blank(), legend.title = element_blank()) +ggtitle("Clustering with 4 groups")

readmission <- readmission %>% mutate(los_age_clust = LOS_Age_Clust)
```

## Task 4 - Select an interaction (5 points)

The following code may help determine if interactions are present. It is best to treat ER as a factor variable for this purpose.

```{r}
#Both variables are factor variables
ggplot(readmission,aes(Gender,fill=factor(Readmission.Status))) + geom_bar(position = "fill") +
   facet_wrap(~Race,ncol=2,scales="free")+scale_y_continuous()

#Look at the change in the percentage of readmitted patients
readmission %>% 
  dplyr::group_by(Race, Gender) %>% 
  dplyr::summarise(percent_readmission = mean(Readmission.Status)) %>% 
  tidyr::spread(Gender, percent_readmission) %>% 
  mutate_if(is.numeric, ~round(.x,2)) %>% 
  make_table()


#One factor variable and one continuous numeric variable
ggplot(readmission,aes(x=factor(Readmission.Status),y=log_riskscore)) + geom_boxplot() +facet_wrap(~factor(ER))
```

## Task 5 - Select a link function (8 points)

The following code runs a GLM using the logit link and all available variables. It assumes that train and test sets have been constructed. Adding an interaction of Gender and Race is included in the code. That is for illustration purposes. The code also produces an ROC curve, a confusion matrix, and calculates AUC.

```{r}
readmission <- readmission %>% select(-DRG.Complication)

#In base R,
#readmission$DRG.Complication <- NULL
#readmission$DRG.Class <- NULL


#Create train and test sets
library(caret)
set.seed(4321)
partition <- createDataPartition(readmission$Readmission.Status, list = FALSE, p = .75) #The partition will stratify using variable 1 from the dataframe
train <- readmission[partition, ]
test <- readmission[-partition, ]

print("TRAIN")
mean(train$Readmission.Status)

print("TEST")
mean(test$Readmission.Status)
```

```{r}
library(pROC)
model <- glm(Readmission.Status ~ . + Gender*Race - los_age_clust, data=train, family = binomial(link="logit"))

summary(model)

preds <- predict(model, newdat=test,type="response")

roc_model <- roc(test$Readmission.Status,preds)

#confusionMatrix(factor(1*(preds>.8)),factor(test$Readmission.Status))
#plot(roc_model)
auc(roc_model)
```

```{r}
model <- glm(Readmission.Status ~ . + Gender*Race - los_age_clust, data=train, family = binomial(link="probit"))

#summary(model)

preds <- predict(model, newdat=test,type="response")

roc_model <- roc(test$Readmission.Status,preds)

#confusionMatrix(factor(1*(preds>.8)),factor(test$Readmission.Status))
#plot(roc_model)
auc(roc_model)
```

```{r}
model <- glm(Readmission.Status ~ . + Gender*Race  - los_age_clust, data=train, family = binomial(link="cauchit"))

#summary(model)

preds <- predict(model, newdat=test,type="response")

roc_model <- roc(test$Readmission.Status,preds)

#confusionMatrix(factor(1*(preds>.8)),factor(test$Readmission.Status))
#plot(roc_model)
auc(roc_model)
```

```{r}
model <- glm(Readmission.Status ~ . + Gender*Race - los_age_clust, data=train, family = binomial(link="cloglog"))

#summary(model)

preds <- predict(model, newdat=test,type="response")

roc_model <- roc(test$Readmission.Status,preds)

#confusionMatrix(factor(1*(preds>.8)),factor(test$Readmission.Status))
#plot(roc_model)
auc(roc_model)
```

## Task 6 -  Decide on the factor variable from Task 3 (5 points)

No code is provided.

```{r}
model <- glm(Readmission.Status ~ . + Gender*Race - log_LOS - Age, data=train, family = binomial(link="logit"))

summary(model)

preds <- predict(model, newdat=test,type="response")

roc_model <- roc(test$Readmission.Status,preds)

#confusionMatrix(factor(1*(preds>.8)),factor(test$Readmission.Status))
#plot(roc_model)
auc(roc_model)
```

## Task 7 - Select features (15 points)

No code is provided.

```{r}
library(caret)

factor_names <- c("DRG.Class","Race") #insert the column names of the variables to be binarized
factor_vars <- readmission[,factor_names]

binarizer <- caret::dummyVars(paste("~", paste(factor_names, collapse = "+")) , data = factor_vars, fullRank = TRUE)
binarized_vars <- data.frame(predict(binarizer, newdata = factor_vars))
head(binarized_vars)

readmission_binarized <- readmission %>% 
  select(-DRG.Class, -Race) %>% 
  mutate(DRG.Class.MED = binarized_vars$DRG.Class.MED,
         DRG.Class.OTHER = binarized_vars$DRG.Class.OTHER,
         DRG.Class.SURG = binarized_vars$DRG.Class.SURG,
         Race.Black = binarized_vars$Race.Black,
         Race.Hispanic = binarized_vars$Race.Hispanic,
         Race.Others = binarized_vars$Race.White)

train_bin <- readmission_binarized[partition,]
test_bin <- readmission_binarized[-partition,]
glimpse(train_bin)
```


```{r}
model <- glm(Readmission.Status ~ . - los_age_clust, data=train_bin, family = binomial(link="logit"))

summary(model)

MASS::stepAIC(model)

preds <- predict(model, newdat=test_bin,type="response")

roc_model <- roc(test$Readmission.Status,preds)

#confusionMatrix(factor(1*(preds>.8)),factor(test$Readmission.Status))

plot(roc_model)
auc(roc_model)
```

```{r}
model <- glm(Readmission.Status ~ Age + log_LOS + log_riskscore + DRG.Class.OTHER + 
    Race.Black + Race.Hispanic,
    data=train_bin, family = binomial(link="logit"))

summary(model)

preds <- predict(model, newdat=test_bin,type="response")

roc_model <- roc(test$Readmission.Status,preds)


plot(roc_model)
auc(roc_model)
```


## Task 8 - Interpret the model (6 points)

No code is provided.

```{r}
model <- glm(Readmission.Status ~ Age + log_LOS + log_riskscore + DRG.Class.OTHER + 
    Race.Black + Race.Hispanic, 
    data = readmission_binarized, 
    family = binomial(link="probit"))

summary(model)
```

```{r}
readmission_binarized %>% summary()
```

```{r}
example1 <- tibble(
  Gender  = "F",
  ER = 0,
  Age = 74,
  log_LOS = 1.65,
  log_riskscore = 0.6,
  los_age_cluster = 2,
  DRG.Class.OTHER = 0,
  DRG.Class.SURG = 1,
  Race.Black = 0,
  Race.Hispanic = 0
)

example2 <- example1 %>% mutate(log_riskscore = 2)
example3 <- example1 %>% mutate(log_LOS = 2)
example4 <- example1 %>% mutate(Age = 80)
example5 <- example1 %>% mutate(Gender = "M")
example6 <- example1 %>% mutate(Race.Black  = 1)
example7 <- example1 %>% mutate(Race.Hispanic  = 1)

preds1 <- predict(model, newdat=example1,type="response")
preds2 <- predict(model, newdat=example2,type="response")
preds3 <- predict(model, newdat=example3,type="response")
preds4 <- predict(model, newdat=example4,type="response")
preds5 <- predict(model, newdat=example5,type="response")
preds6 <- predict(model, newdat=example6,type="response")
preds7 <- predict(model, newdat=example7,type="response")

example1 %>% 
  rbind(example2) %>% 
  rbind(example3) %>% 
  rbind(example4) %>% 
  rbind(example5) %>% 
  rbind(example6) %>% 
  rbind(example7) %>% 
  mutate(y = c(preds1, preds2, preds3, preds4,preds5, preds6, preds7)) %>% 
  mutate(patientid = row_number(), y = round(y,3)) %>% 
  select(patientid, y)
```


## Task 9 - Set the cutoff (9 points)

The following code calculates the cost using a cutoff of 0.075. It assumes the final model constructed on the full dataset is called glm_full and the final dataset is readmit.

```{r}
readmission$Readmission.Status %>% sum()*25

pred_full <- predict(model,type="response")

cutoff <- 0.1

pred_readmit <- 1*(pred_full > cutoff)

no_intervention_cost <- 25*sum(readmission$Readmission.Status == 1)
full_intervention_cost <- 2*nrow(readmission)
no_intervention_cost
full_intervention_cost

library(e1071)
cm <- confusionMatrix(factor(1*(pred_full>cutoff)),factor(readmission$Readmission.Status))$table
cm

modified_cost <- cm[2,1]*2+cm[2,2]*2+cm[1,2]*25
modified_cost


```

```{r}
cutoff_values <- c(1, 0.4, 0.3, 0.2, 0.1, 0.09, 0.08, 0.075, 0.07, 0.05, 0)

get_modified_cost <- function(input_cutoff){
  pred_readmit <- 1*(pred_full > input_cutoff)
  cm <- confusionMatrix(factor(pred_readmit),factor(readmission$Readmission.Status))
  modified_cost <- cm$table[2,1]*2+cm$table[2,2]*2+cm$table[1,2]*25
  modified_cost
}
get_modified_cost(0.2)


tibble(cutoff = cutoff_values,
       cost = sapply(cutoff_values, get_modified_cost)) 
  
```

```{r}
cutoff <- 0.08
pred_readmit <- 1*(pred_full > cutoff)
cm <- confusionMatrix(factor(pred_readmit),factor(readmission$Readmission.Status))
cm$table
```

## Task 10 – Consider alternative models and model construction techniques (12 points)


## Task 11 – Executive summary (20 points)
