---
title: "Patient Length of Stay -  Solution"
subtitle: "ExamPA.net Solution to SOA PA 6/18/20"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    theme: cerulean
    toc: true
    code_folding: show
    toc_float: true
---

For your convenience, this solution document has been formatted to be easy to read.  When taking this exam, candidates are not expected to make their `.Rmd` documents look this way.  The `make_table` function is included only to improve your reading experience.  The SOA's instructions from the Project Statement say only that

>At a minimum you must submit your completed report template and an Rmd file that supports your work.  Graders expect that your Rmd code can be run from beginning to end.  The code snippets provided should either be commented out or adapted for execution.  Ensure that it is clear where in the code each of the Tasks is addressed.  Documentation of your thought process and conclusion for each Task should be completely contained within your Word report.  The Rmd code should be clear, may contain commentary, and support your work.

---
Your assistant has provided the following code to load the dataset and provide summary values.

```{r setup, warning = F, message = F}
#NO CHANGES NEEDED

# Load packages and data

# Load packages
library(tidyverse)
library(kableExtra) #If needed, install this package.  It's just used for making tables.
library(ExamPAData)
library(plyr)
library(dplyr)

theme_set(theme_minimal())

#function to create html tables
make_table <- function(df, caption = ""){
  df %>% 
    kable(caption = caption) %>% 
    kable_styling(bootstrap_options = "striped", full_width = F)}

#global chunk options
knitr::opts_chunk$set(warning = F, message = F, echo = T, fig.width = 14, fig.height = 8)

# Load data
data.all <- patient_length_of_stay
summary(data.all)
str(data.all)

data.all %>% mutate_if(is.character, as.factor) %>% summary()
```

# Task 1

Edit the data for missing and invalid data.

Generic code has been provided to edit the data. You can modify and use some or all of this code to edit the data.

Convert numeric variable type to factor variable type

```{r warning = F, message = F}
# Replace VARIABLE with the variable name of the variable to be changed.

#data.all$VARIABLE <- as.factor(data.all$VARIABLE) # Replace VARIABLE twice
#str(data.all)

data.all$age <- as.factor(data.all$age) # Replace VARIABLE twice
data.all$admit_type_id <- as.factor(data.all$admit_type_id) # Replace VARIABLE twice

data.all <- data.all %>% filter(gender != "Unknown/Invalid:")

data.all <- data.all %>% filter(gender %in% c("Male", "Female"))

#gender_missing <- data.all$gender == "Unknown/Invalid:"
#data.all <- data.all[-gender_missing,]

data.all <- data.all %>% mutate(race = ifelse(race == "?", "Other", race))

#data.all$race <- ifelse(data.all$race == "?", "Other", data.all$race)

data.all <- data.all %>% select(-weight)
data.all$weight <- NULL # Replace VARIABLE

data.all <- data.all %>% mutate(race = ifelse(race == "Asian", "Other", race))
data.all$race <- ifelse(data.all$race=="Asian", "Other", data.all$race)

library(forcats)
data.all <- data.all <- data.all %>% mutate_if(is.factor, fct_infreq)

data.all <- data.all <- data.all %>% mutate_if(is.character, fct_infreq)

summary(data.all)

data.all %>% dplyr::count(num_procs)
data.all %>% dplyr::count(num_meds)
data.all %>% dplyr::count(num_ip)
data.all %>% dplyr::count(num_diags)

summary(data.all)
```

Remove variables / columns

```{r eval = F}
# Replace VARIABLE with the variable name of the variable to be removed.

data.all$VARIABLE <- NULL # Replace VARIABLE
head(data.all)
```

Remove observations / rows

```{r eval = F}

# Replace VARIABLE with the variable name with values to be removed.
# Replace VALUE with the value of the variable to be removed.

data.all <- subset(data.all, VARIABLE != "VALUE") # Replace VARIABLE and VALUE
# Value is in quotation marks for factor variables. No quotation marks for numeric variables.

data.all$VARIABLE <- droplevels(data.all$VARIABLE) # If VARIABLE is a factor variable this must be run else the dropped factor level will be retained, but with zero observations.

summary(data.all$VARIABLE)
```

Combine variable levels

```{r eval = F}
# Replace VARIABLE with the variable name to have reduced number of levels.
# Replace LEVELs with new level names.

print("Data Before Combine Levels")
table(data.all$VARIABLE) # Replace VARIABLE

# Combine levels of VARIABLE by mapping one level to another level

var.levels <- levels(data.all$VARIABLE) # Replace VARIABLE
data.all$VARIABLE <- mapvalues(
  data.all$VARIABLE, var.levels,
  c("LEVEL1", "LEVEL2", "LEVEL3", "LEVEL4", "LEVEL5", "LEVEL6")
) # Replace VARIABLE twice and replace LEVELs with the new names.

print("Data After Combine Levels")
table(data.all$VARIABLE) # Replace VARIABLE

rm(var.levels)
```

Relevel factor variables.

Change list of factor variables if you removed one of the variables or if you converted a numeric variable to a factor variable.

```{r eval = F}

vars <- c("gender", "age", "race", "metformin", "insulin", "readmitted") # Change list of factor variables to relevel as needed

for (i in vars) {
  table <- as.data.frame(table(data.all[, i]))
  max <- which.max(table[, 2])
  level.name <- as.character(table[max, 1])
  data.all[, i] <- relevel(data.all[, i], ref = level.name)
}
```

# Task 2

Explore the data. 

Examine the variables and their relationships to the target variable.

Code has been provided to explore the data. You can modify and use some or all of this code to explore the data.

Descriptive statistics

```{r}

summary(data.all)
summary(data.all$days)
```
Bar charts

```{r warning = F, message = F}
library(ggplot2)
vars <- colnames(data.all)

for (i in vars) {
  plot <- ggplot(data.all, aes(x = unlist(data.all[, i]))) +
    geom_bar() +
    labs(x = i) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  print(plot)
}

rm(i, vars, plot)
```

Target variable means by predictor variables.

```{r}
# This chunk provides means of the target variable split by predictors.
vars <- colnames(data.all)

for (i in vars) {
  x <- data.all %>%
    group_by_(i) %>%
    dplyr::summarise(
      mean = mean(days),
      median = median(days),
      n = dplyr::n()
    )

  print(x)
}

rm(i, vars, x)
```

Correlations of the target variable to numeric variables and correlation matrix.

```{r}

# Calculate the correlation matrix for numeric variables
cor.matrix <- cor(data.all[, sapply(data.all, is.numeric)])

print("Correlation Matrix")
cor.matrix
```

Split histograms and boxplots of target by factor variables. 

Copy and add code for other factors as needed.

```{r warning = F, message = F}

data.all %>% str()

library(gridExtra)

# Explore target days vs. gender.

# Split histogram
p1 <- ggplot(data.all, aes(
  x = days,
  group = readmitted   , fill = readmitted   , y = ..density..
)) +
  geom_histogram(position = "dodge", binwidth = 1) # Replace gender twice for other variables

# Boxplot
p2 <- ggplot(data = data.all, aes(
  x = readmitted   , y = days,
  fill = readmitted   
)) +
  geom_boxplot(alpha = 0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) # Replace gender twice for other variables

grid.arrange(p1, p2, ncol = 2)
```

```{r warning = F, message = F}
library(dplyr)
library(ggplot2)
library(tidyverse)
data.all %>% 
  ggplot(aes(days, num_meds )) + 
  geom_point(size = 1.5, alpha = 0.2, color = "blue") + 
  theme_minimal() + 
  ggtitle("Number of Days vs. Number of Medications") + 
  xlab("Number of days") + 
  ylab("Number of Medications")
```


# Task 3

Consider two data issues.

No code provided.

# Task 4

Write a data summary for your actuarial manager.

No code provided.

# Task 5

Perform Principal Components Analysis (PCA).

The following chunk performs PCA on numeric variables. 

```{r}
# Select only the variables used for PCA. Do not change this list.
data.pca <- data.all[, c("num_procs", "num_meds", "num_ip", "num_diags")]

# Run PCA on the numeric variables. Variables are centered and scaled.
pca <- prcomp(data.pca, center = TRUE, scale. = TRUE)
summary(pca)
pca$rotation
plot(pca, type = "line")
```

The following chunk constructs a new feature using the first principal component and attaches it to the data frame.

```{r}

# Center and scale the variables
data.pca.std <- as.data.frame(scale(data.pca))
head(data.pca.std)

# Add the first principal component to the data frame.
data.all$PC1 <- pca$x[, 1]
str(data.all$PC1)
```

Split the data into training and test data sets.

```{r warning = F, message = F}
# Split the data into training and test data sets.

library(caret)
set.seed(100)
train_ind <- createDataPartition(data.all$days, p = 0.7, list = FALSE)
data.train <- data.all[train_ind, ]
data.test <- data.all[-train_ind, ]

print("TRAIN")
mean(data.train$days)

print("TEST")
mean(data.test$days)

print("ALL")
mean(data.all$days)


rm(train_ind)
```

# Task 6

Construct a decision tree.

Construct an unpruned regression tree. Do not use PC1. Do not change the code.

```{r warning = F, message = F}
library(rpart)
library(rpart.plot)
set.seed(555)

# Fit the model
tree1 <- rpart(days ~ . - PC1,
  data = data.train,
  method = "anova",
  control = rpart.control(cp = 0.001, minbucket = 20)
)

tree1

# Plot the tree
rpart.plot(tree1)

# Predict on training data
tree1.pred.train <- predict(tree1, data.train, type = "vector")

# Calculate the Pearson goodness-of-fit statistic on training data
sum((data.train$days - tree1.pred.train)^2 / tree1.pred.train) / nrow(data.train)

# Predict on test data
tree1.pred.test <- predict(tree1, data.test, type = "vector")

# Calculate the Pearson goodness-of-fit statistic on test data
sum((data.test$days - tree1.pred.test)^2 / tree1.pred.test) / nrow(data.test)
```

Use cost-complexity pruning to produce a reduced tree.

```{r}
# Display the complexity parameter table and plot for tree1.
tree1$cptable
plotcp(tree1)
```

```{r}
# Prune the tree by replacing XX with the complexity parameter that will result in eight leaves.
# If eight is not a possible option, select the largest number less than eight that is possible.
tree2 <- prune(tree1, cp = 0.003568539) # Replace XX.

tree2

# Plot the tree
rpart.plot(tree2)

# Predict on training data
tree2.pred.train <- predict(tree2, data.train, type = "vector")

# Calculate the Pearson goodness-of-fit statistic on training data
sum((data.train$days - tree2.pred.train)^2 / tree2.pred.train) / nrow(data.train)

# Predict on test data
tree2.pred.test <- predict(tree2, data.test, type = "vector")

# Calculate the Pearson goodness-of-fit statistic on test data
sum((data.test$days - tree2.pred.test)^2 / tree2.pred.test) / nrow(data.test)
```

# Task 7

Construct a generalized linear model (GLM).

Fit a GLM with all original variables included but excluding the PCA variable. Do not change the Poisson distribution or log link function.

```{r}
# Fit the model
glm1 <- glm(days ~ . - PC1,
  data = data.train,
  family = poisson(link = "log") # Do not change.
)

summary(glm1)

# Predict on training data
glm1.pred.train <- predict(glm1, data.train, type = "response")

# Calculate the Pearson goodness-of-fit statistic on training data
sum((data.train$days - glm1.pred.train)^2 / glm1.pred.train) / nrow(data.train)

# Predict on test data
glm1.pred.test <- predict(glm1, data.test, type = "response")

# Calculate the Pearson goodness-of-fit statistic on test data
sum((data.test$days - glm1.pred.test)^2 / glm1.pred.test) / nrow(data.test)
```

Fit a GLM with the PCA variable created in Task 5 included (and without the numeric variables used to produce the PCA variable). Do not change the Poisson distribution or log link function.

```{r}
# Fit the model
glm2 <- glm(days ~ . - num_procs - num_meds - num_ip - num_diags,
  data = data.train,
  family = poisson(link = "log") # Do not change.
)

summary(glm2)

# Predict on training data
glm2.pred.train <- predict(glm2, data.train, type = "response")

# Calculate the Pearson goodness-of-fit statistic on training data
sum((data.train$days - glm2.pred.train)^2 / glm2.pred.train) / nrow(data.train)

# Predict on test data
glm2.pred.test <- predict(glm2, data.test, type = "response")

# Calculate the Pearson goodness-of-fit statistic on test data
sum((data.test$days - glm2.pred.test)^2 / glm2.pred.test) / nrow(data.test)
```

# Task 8

Perform feature selection with lasso regression.

Use cross-validation to determine appropriate level of lambda for lasso regression.

```{r warning = F, message = F}
library(glmnet)

# Format data as matrices (necessary for glmnet). Uncomment two items that reflect your decision from Task 7.

lasso.mat.train <- model.matrix(days ~ . - PC1, data.train)
lasso.mat.test <- model.matrix(days ~ . - PC1, data.test)
# lasso.mat.train <- model.matrix(days ~ . - num_procs - num_meds - num_ip - num_diags, data.train)
# lasso.mat.test <- model.matrix(days ~ . - num_procs - num_meds - num_ip - num_diags, data.test)

set.seed(789)

lasso.cv <- cv.glmnet(
  x = lasso.mat.train,
  y = data.train$days,
  family = "poisson", # Do not change.
  alpha = 1 # alpha = 1 for lasso
)
```

Use the cross-validation results to run the final lasso regression model.

```{r}
# Fit the model
lasso <- glmnet(
  x = lasso.mat.train,
  y = data.train$days,
  family = "poisson", # Do not change.
  lambda = lasso.cv$lambda.1se,
  alpha = 1 # alpha = 1 for lasso
)

# List variables
lasso$beta

# Predict on training data
lasso.pred.train <- predict(lasso, lasso.mat.train, type = "response")

# Calculate the Pearson goodness-of-fit statistic on training data
sum((data.train$days - lasso.pred.train)^2 / lasso.pred.train) / nrow(data.train)

# Predict on test data
lasso.pred.test <- predict(lasso, lasso.mat.test, type = "response")

# Calculate the Pearson goodness-of-fit statistic on test data
sum((data.test$days - lasso.pred.test)^2 / lasso.pred.test) / nrow(data.test)
```

# Task 9

No code provided.

# Task 10

No code provided.

# Task 11

Interpret the model for the client.

Copy the GLM code for the recommended model from Task 7 and run on the full dataset to interpret coefficients.

```{r}
glmall <- glm(days ~ . - PC1,
  data = data.all,
  family = poisson(link = "log") # Do not change.
)

summary(glmall)
summary(data.all)
```


# Task 12

No code provided.
