---
title: 'Titanic Template'
subtitle: "ExamPA.net Practice Exam"
author: "Sam Castillo"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    theme: cerulean
    toc: true
    code_folding: hide
    toc_float: true
---

Your assistant has provided code throughout this document.  The structure of the code can always be treated as being correct, however, there is no gaurantee that the parameter choices are best.

```{r, message = FALSE}
#NO CHANGES NEEDED
set.seed(1234)
#install these packages
#install.packages(c("car", "fastDummies", "ROCR"))
library(caret) 
library(randomForest) 
library(rpart)
library(rpart.plot)
library(car)
library(e1071)
library(broom)
library(fastDummies)
library(caret)
library(ROCR)
library(tidyverse)
library(glmnet)
library(dplyr)
library(scales)
library(ExamPAData)
theme_set(theme_minimal())
```

```{r}
#NO CHANGES NEEDED
#convert factor levels to those with the most observations
df <- exam_pa_titanic %>% 
  mutate_at(.vars = c("sex", "embarked"), .funs = fct_infreq)
```

## Task 1 - Explore the data


```{r}
#NO CHANGES NEEDED
df %>% count(survived) %>% mutate(percent_of_passengers = percent(n/sum(n))) 
```

```{r}
#NO CHANGES NEEDED
df %>% 
  group_by(pclass) %>% 
  summarise(n = n(),
            percent_survived = percent(sum(survived==1)/n()))

df %>% 
  group_by(sex) %>% 
  summarise(n = n(),
          percent_survived = percent(sum(survived==1)/n()))

df %>% 
  group_by(embarked) %>% 
  summarise(n = n(),
           percent_survived = percent(sum(survived==1)/n()))
```

```{r fig.width=6, fig.height=4}
summary(df$fare)
df %>% 
  ggplot(aes(as.factor(survived), fare)) + 
  geom_boxplot()

df <- df %>% mutate(fare = log(fare))

#In Base R
#df$log_fare = log(df$fare)
#df$fare <- NULL

df %>% ggplot(aes(fare)) + geom_histogram()

summary(df$age)
df %>% 
  ggplot(aes(as.factor(survived), age)) + 
  geom_boxplot()
df %>% ggplot(aes(age)) + geom_histogram()
```


## Task 2 - Create a new variable called "title"

Your assistant has provided the following code to create the new "title" variable.

```{r, message=FALSE, warning=FALSE}
##Lets extract the title and check if we have predictive power in that
df <- df %>% mutate(title = gsub("^.*, (.*?)\\..*$", "\\1", name))

df %>% count(title)

df <- df %>% mutate(
  title = case_when(
    title %in% c("Miss","Mlle", "MS", "Ms") ~ "Miss", #NEEDS TO BE UPDATED
    title == "Mme" ~ "Mrs", 
    title %in% c("Officer") ~ "Officer",
    T ~ title
  )%>% fct_infreq()
) 
df %>% count(title) 
```

This code is the same as above except it uses Base R instead of dplyr.

```{r, eval = F, message=FALSE, warning=FALSE}
##Lets extract the title and check if we have predictive power in that
df$title <- gsub("^.*, (.*?)\\..*$", "\\1", df$name)

table(df$title)

#This code simplies the title
#For example, passengers which have "Miss", "Mlle", "MS", or "Ms" in their name get the title "Miss"
df$title <- ifelse(df$title %in% c("Miss","Mlle", "MS", "Ms"), yes = "Miss", no = df$title)
df$title <- ifelse(df$title == "Mme", yes = "Mrs", no = df$title)
df$title <- ifelse(df$title %in% c("Officer", "Capt"), yes = "Officer", no = df$title)

table(df$title)

#Set the base factor level to the title with the most observations
df$title <- fct_infreq(df$title)
```

## Task 3 - Create a new variable called "family_size"

```{r}
#example with dplyr
df <- df %>% mutate(family_size = 1)

#example in base R
df$family_size = 1
```

## Task 4

```{r}
#NO CHANGES NEEDEDset.seed(1234)
cluster_data <- df %>% 
  select(age, fare,  pclass, family_size) %>% 
  mutate(fare = exp(fare))#Note: if you applied a log transform to fare, undo this before clustering

run_kmeans <- function(centers, nstart){
  kmeans = kmeans(cluster_data, centers = centers, nstart = nstart)

  df <- cluster_data %>% mutate(
    cluster = as.factor(kmeans$cluster)
  ) 
  
  counts <- df %>% count(cluster)
  list("kmeans" = kmeans, "counts" = counts)
}
```

```{r}
#Run KMeans
#Choose values for centers and nstart
run1 <- run_kmeans(centers = 10, nstart = 3)
run1$counts

#select the kmeans run to use
#select the clusters which you are counting as outliers
#i.e., if clusters 4 and 5 were the "outliers", say outlier_clusters <- c(4,5)

selected_run <- run1
outlier_clusters <- c(2)

#add selected cluster assignment to the data
df <- df %>% mutate(
  cluster = selected_run$kmeans$cluster %>% as.factor(),
  outlier_flag = ifelse(cluster %in% outlier_clusters, 1, 0) %>% as.factor() %>% fct_infreq()
) 
```

```{r}
#NO CHANGES NEEDED
#Cluster Assignments
p1 <- df %>% ggplot(aes(age,fare, color = cluster)) + geom_point() + theme(legend.position = "none")
p2 <- df %>% ggplot(aes(pclass,family_size, color = cluster)) + geom_point()+ theme(legend.position = "none")
p3 <- df %>% ggplot(aes(fare,family_size, color = cluster)) + geom_point()+ theme(legend.position = "none")
p4 <- df %>% ggplot(aes(age,pclass, color = cluster)) + geom_point()+ theme(legend.position = "none")
gridExtra::grid.arrange(p1,p2,p3, p4)
df <- df %>% select(-cluster)
```

```{r}
#NO CHANGES NEEDED
#Probable Outliers
p1 <- df %>% ggplot(aes(age,fare, color = outlier_flag)) + geom_point() + theme(legend.position = "none")
p2 <- df %>% ggplot(aes(pclass,family_size, color = outlier_flag)) + geom_point()+ theme(legend.position = "none")
p3 <- df %>% ggplot(aes(fare,family_size, color = outlier_flag)) + geom_point()+ theme(legend.position = "none")
p4 <- df %>% ggplot(aes(age,pclass, color = outlier_flag)) + geom_point()+ theme(legend.position = "none")
gridExtra::grid.arrange(p1,p2,p3, p4)
```

## Task 5

```{r}
#SELECT WHICH VARIABLES YOU WANT TO REMOVE

#your assistent dropped passenger id
#if you decide to drop other variables, add them to this code
#example: drop passenger id and fare
#df <- df %>% select(-passengerid, -fare)
df <- df %>% select(-passengerid)
```

```{r}
#NO CHANGES NEEDED
#create dummy columns and split into train and test sets
df_factors <- df %>% select(pclass, sex, embarked, title)
df_dummies = df %>% 
  dummy_cols(remove_first_dummy = T) %>%
  select(-sex, -title, -embarked)
  
df_dummies %>% dim()
df <- df_dummies

train_index <- createDataPartition(df$survived, p = 0.75, list = F) %>% as.numeric()
train <- df %>% dplyr::slice(train_index)
test <- df %>% dplyr::slice(-train_index)
train %>% summarise(train_percent_survived = sum(survived==1)/n())
test %>% summarise(test_percent_survived = sum(survived==1)/n())
dim(train)
```

## Task 6

```{r, message=FALSE, warning=FALSE}
rf <- randomForest(survived~.,
                     data = train,
                     ntree = 3,
                     mtry = 300,
                  importance = T)
plot(rf)
predictions <- as.vector(predict(rf, test))
pred <- prediction(predictions, test$survived)

accuracy <- sum((predict(rf, test) > 0.5) == test$survived)/nrow(test)

perf_measures <- performance(pred, "auc")
auc <- perf_measures@y.values[[1]]

print("Accuracy")
accuracy

print("AUC")
auc
#note: ignore the error "The response has five or fewer unique values.  Are you sure you want to do regression?".  This is due to a bug in the randomForest library.
```


## Task 7

```{r}
importance <- varImp(rf)
rf_results <- tibble(variable = rownames(importance),
       importance = importance$Overall) 
rf_results
```


## Task 8

```{r, message=FALSE, warning=FALSE}
#choose variables to use 
logit_vars <- c("survived", "pclass", "age", "fare", "family_size", "outlier_flag", "sibsp", "parch")

logit_train <- train %>% select(logit_vars)
logit_test <- test %>% select(logit_vars)

model <- glm(survived ~ ., 
            family = binomial(link="logit"), 
            data = logit_train)
summary(model)
model %>% tidy() %>% select(term, estimate) 
```

## Task 9

```{r}
#NO CHANGES NEEDED
#this code gets the mean (for continuous) and mode (for categorical) variables
continuous_vars <- c("fare", "family_size", "sibsp")
train %>% select(continuous_vars) %>% summarise_all(~round(mean(.x),1))

categorical_vars <- c("pclass", "outlier_flag")
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
train %>% select(categorical_vars) %>% summarise_all(getmode)
```

```{r}
#check that the Excel calculation of the survival probability matches the R calculation
#create a fake passenger with age 10
person_age_10 <- tibble(age = 10, #Update these values to be the median/mode of each variable
                        fare = 3,
                        family_size = 2.1,
                        sibsp = 0.7,
                        pclass = 3,
                        outlier_flag = 0)
```

```{r}
#NO CHANGES NEEDED
#check your work here
predict(model, person_age_10, type = "response")
person_age_30 = person_age_10 %>% mutate(age = 30)
predict(model, person_age_30, type = "response")
person_age_60 = person_age_10 %>% mutate(age = 60)
predict(model, person_age_60, type = "response")
```

## Task 10

```{r}
#NO CHANGES NEEDED
predictions <- as.vector(predict(model, test))
pred <- prediction(predictions, test$survived)
perf_measures <- performance(pred, "auc")
auc <- perf_measures@y.values[[1]]

print("Accuracy")
sum((predict(model, test) > 0.5) == test$survived)/nrow(test)

print("AUC")
auc
```

## Task 11

```{r message=F}
selected_alpha <- 0.5

x <- model.matrix(survived ~ .^2, data = logit_train )
colnames(x)

lasso_cv <- cv.glmnet(x = x, 
            y = train$survived,
            family = "binomial",
            alpha = selected_alpha) 
```

## Task 12

```{r}
plot(lasso_cv)
```

```{r}
model <- glmnet(x = x, 
            y = logit_train$survived,
            family = "binomial", 
            alpha = selected_alpha) 

plot(model, label = T, xvar = "dev")
```

```{r}
#make selections
selected_lambda <- 0.01
selected_lambda <- exp(-3)


model <- glmnet(x = x, 
            y = logit_train$survived,
            family = "binomial", 
            lambda = selected_lambda,
            alpha = selected_alpha) 

model$beta
```

## Task 13

```{r}
#NO CHANGES NEEDED
x_test <- model.matrix(survived ~ .^2, data = logit_test)

predictions <- as.vector(predict(model, x_test))
pred <- prediction(predictions, test$survived)
perf_measures <- performance(pred, "auc")
auc <- perf_measures@y.values[[1]]

print("Accuracy")
sum((predictions > 0.1) == test$survived)/nrow(test)
print("AUC")
auc
```

